---
title: Analysis of Micromobility Dock Deployment Strategies
authors:
  - name: Joshua Sweren
    affiliation: Georgetown University
    roles: writing
    corresponding: true
format:
  html:
    embed-resources: true
    code-fold: true
    echo: true
execute:
  warning: false
  message: false
toc: true
---

```{r, include=FALSE}
#install.packages(c("osmdata", "sf", "dplyr", "ggplot2", "rnaturalearth"))
library(osmdata)
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(osmextract)
library(tigris)
library(mapview)
library(lehdr)
library(viridis)
library(viridisLite)
library(spatstat)
library(units)
library(spdep)
library(terra)
library(raster)
library(tictoc)
library(tidycensus)
library(leaflet)
```

Github repository: <https://github.com/jsweren1/gis-final-project>

## Introduction

Over the last decade, bikeshare systems have become increasingly prevalent in cities around the world in recent years, offering residents and visitors a convenient, low-cost, and environmentally friendly transportation option. However, the spatial deployment of these systems has significant implications over which communities benefit the most and the extent to which they can be of service. This project will analyze the geographic distribution of micromobility stations in several metropolitan areas and evaluate how access to these low-cost transportation options varies across neighborhoods.

As of October 2025, there are five metropolitan areas in the U.S. whose docked bikeshare ridership far exceeds the rest: New York City, Boston, San Francisco, Chicago, and Washington, DC. These cities’ bikeshare systems each boasted ridership of over 400,000 in October 2025, while the sixth-place city registered under 100,000.[^1] These numbers are similarly disparate in other months. Thus, this project will focus on five cities with proven commitment to developing a system with geographic breadth. Unsurprisingly, these cities also have existing robust public transit networks, so this project will not only look at spatial deployment in relation to population, but also in relation to existing amenities for certain residents. Using OpenStreetMap data to locate bikeshare docks and other transit hubs, along with U.S. Census data to map population density, we will compare how bikeshare has been spatially deployed in its most prominent cases.

[^1]: “Docked bikeshare trips – October 2025,” Bureau of Transportation Statistics, https://www.bts.gov/newsroom/docked-bikeshare-trips-october-2025 (accessed Dec. 15, 2025).

## Literature Review

The following articles were provided important context on what research has already been done regarding this topic:

1.  *Optimal locations for bikeshare stations: A new GIS based spatial approach*[^2]
2.  *Using a GIS-based spatial approach to determine the optimal locations of bikeshare stations: The case of Washington D.C*[^3]

[^2]: S. Banerjee, Md. M. Kabir, N. K. Khadem, and C. Chavis, “Optimal locations for bikeshare stations: A new GIS based spatial approach,” *Transportation Research Interdisciplinary Perspectives*, vol. 4, p. 100101, Mar. 2020. doi:10.1016/j.trip.2020.100101

[^3]: Z. Dehdari Ebrahimi, M. Momenitabar, A. A. Nasri, and J. Mattson, “Using a GIS-based spatial approach to determine the optimal locations of bikeshare stations: The case of Washington D.C,” *Transport Policy*, vol. 127, pp. 48–60, Oct. 2022. doi:10.1016/j.tranpol.2022.08.008

These articles analyzed bikeshare dock location strategies in Baltimore and Washington, D.C., respectively. The authors studying Baltimore focused mainly on optimizing usage of these docks and understanding what factors lead to greater ridership. They found that bikeshare usage depended heavily on the presence of neighboring docks, meaning clustered docks tended to benefit people more than isolated ones. In the case of Baltimore, the bikeshare system is significantly less thorough and mature than the five cities we will be studying, so the effects of isolated bikeshare docks were perhaps more apparent than we would find. This indicates that it will be important to study the clustering of docks as a univariate test.

The authors studying Washington, D.C. also focused on optimal locations, mainly as they relate to proximity to bus stops and train stations. For cities with existing public transit infrastructure, bikeshare can be seen as a supplement to heavily used services, not only as its own form of transportation. However, this study also emphasized the importance of bikeshare in "low-density residential neighborhoods where transit network coverage might be low". This highlights a need for us to look at spatial equity in micromobility infrastructure, as bikeshare systems can provide greater accessibility to cheaper transit options.

## Methodology

Each city will be analyzed via the following sets of hypotheses:

1.  Docks in relation to population density:

    -   **Null** **hypothesis**: There is no relationship between the spatial distribution of bikeshare docks and population density. Dock locations occur independently of areas with higher or lower population density.
    -   **Alternative** **hypothesis**: There is a relationship between the spatial distribution of Capital Bikeshare docks and population density. Dock locations are more likely to occur in areas with higher population density.

2.  Docks in relation to other public transit options:

    -   **Null hypothesis**: There is no relationship between the spatial distribution of bikeshare docks and the presence of other public transit nodes.

    -   **Alternative hypothesis:** There is a relationship between the spatial distribution of bikeshare docks and the presence of other public transit nodes. Dock locations are more likely to be placed in areas with existing public transit infrastructure.

The first steps towards testing these hypotheses will be to gather bikeshare dock and train station locations from OpenStreetMap via the `osm` package in R and population density data from the U.S. Census via the `tidycensus` package. To create a geographic region for each city, we will create a concave hull around all of the bikeshare docks. Since metropolitan areas differ in the distance from the city's center that bikeshare systems reach, this will allow us to only consider areas within the chosen scope of each system, and how docks have been deployed within those limitations.

The following information will then be gathered to test the hypotheses:

-   Moran's I to generally understand clustering of bike docks.

-   Regression testing results to analyze the relationship between dock locations, population density, and public transit options.

-   Monte Carlo simulations to compare actual results to randomized data.

## Exploratory Data Analysis

Prior to performing any hypothesis tests, we will start by plotting all bikeshare docks from each metropolitan area, along with a reasonable geographic region to surround them. The chosen methodology for selecting a region is to take a concave hull using `spdep` in R, with `0.2` as the argument for this function. This was chosen so that each metropolitan area's docks would fit in a contiguous region that accurately represents the extent of the system's reach, while limiting the amount of uninhabitable space that would be captured if we used a convex hull. Padding was added proportionally to the overall extent of each bikeshare system, simply to avoid docks being directly on the edge of our regions.

The plots below visualize bikeshare docks, overlaid over these chosen regions for our five cities of interest.

### Plotting the Docks

::: panel-tabset
### Washington, DC

```{r}
dc_bb <- c(xmin = -77.45, ymin = 38.70, xmax = -76.80, ymax = 39.10)
docks_raw_dc <- opq(bbox = dc_bb) |>
  add_osm_feature(key = "amenity", value = "bicycle_rental") |>
  add_osm_feature(key = "network", value = "Capital Bikeshare") |>
  osmdata_sf()

docks_dc <- docks_raw_dc$osm_points |>
  st_as_sf() |>
  st_transform(3857) %>%
  filter(!is.na(network))
```

```{r}
hull_dc <- docks_dc |> 
    st_union() |> 
    st_concave_hull(.2)
hull_buffered_dc <- st_buffer(hull_dc, dist = 2500)
hull_buffered_wgs84_dc <- st_transform(hull_buffered_dc, 4326)
docks_dc <- st_intersection(
    st_transform(docks_dc, 4326), 
    hull_buffered_wgs84_dc)
```

```{r}
mapview(hull_buffered_wgs84_dc,
        col.regions = "lightblue",
        alpha.regions = 0.3) +
mapview(docks_dc, col.regions = "red",
  layer.name = "Bikeshare Docks")
```

### New York, NY

```{r}
nyc_bb <- c(xmin = -74.15, ymin = 40.55, xmax = -73.70, ymax = 40.95)
docks_raw_nyc <- opq(bbox = nyc_bb) |>
  add_osm_feature(key = "amenity", value = "bicycle_rental") |>
  add_osm_feature(key = "network", value = "Citi Bike") |>
  osmdata_sf()

docks_nyc <- docks_raw_nyc$osm_points |>
  st_as_sf() |>
  st_transform(3857) %>%
  filter(!is.na(network))
```

```{r}
hull_nyc <- docks_nyc |> 
  st_union() |> 
  st_concave_hull(.2)
hull_buffered_nyc <- st_buffer(hull_nyc, dist = 500)
hull_buffered_wgs84_nyc <- st_transform(hull_buffered_nyc, 4326)
docks_nyc <- st_intersection(
  st_transform(docks_nyc, 4326),
  hull_buffered_wgs84_nyc
)
```

```{r}
mapview(hull_buffered_wgs84_nyc, col.regions = "lightblue", alpha.regions = 0.3) +
  mapview(docks_nyc, col.regions = "red",
  layer.name = "Bikeshare Docks")
```

### Boston, MA

```{r}
bos_bb <- c(-71.20, 42.20, -70.90, 42.45)
docks_raw_bos <- opq(bbox = bos_bb) |>
  add_osm_feature(key = "amenity", value = "bicycle_rental") |>
  add_osm_feature(key = "network", value = "Bluebikes") |>
  osmdata_sf()

docks_bos <- docks_raw_bos$osm_points |>
  st_as_sf() |>
  st_transform(3857) %>%
  filter(!is.na(network))
```

```{r}
hull_bos <- docks_bos |> 
  st_union() |> 
  st_concave_hull(.2)
hull_buffered_bos <- st_buffer(hull_bos, dist = 1000)
hull_buffered_wgs84_bos <- st_transform(hull_buffered_bos, 4326)
docks_bos <- st_intersection(
  st_transform(docks_bos, 4326),
  hull_buffered_wgs84_bos
)
```

```{r}
mapview(hull_buffered_wgs84_bos, col.regions = "lightblue", alpha.regions = 0.3) +
  mapview(docks_bos, col.regions = "red",
  layer.name = "Bikeshare Docks")
```

### San Francisco, CA

```{r}
sf_bb <- c(-122.55, 37.70, -122.35, 37.85)
docks_raw_sf <- opq(bbox = sf_bb) |>
  add_osm_feature(key = "amenity", value = "bicycle_rental") |>
  add_osm_feature(key = "network", value = "Bay Wheels") |>
  osmdata_sf()

docks_sf <- docks_raw_sf$osm_points |>
  st_as_sf() |>
  st_transform(3857) %>%
  filter(!is.na(network))
```

```{r}
hull_sf <- docks_sf |> 
  st_union() |> 
  st_concave_hull(.2)
hull_buffered_sf <- st_buffer(hull_sf, dist = 1000)
hull_buffered_wgs84_sf <- st_transform(hull_buffered_sf, 4326)
docks_sf <- st_intersection(
  st_transform(docks_sf, 4326),
  hull_buffered_wgs84_sf
)
```

```{r}
mapview(hull_buffered_wgs84_sf, col.regions = "lightblue", alpha.regions = 0.3) +
  mapview(docks_sf, col.regions = "red",
  layer.name = "Bikeshare Docks")
```

### Chicago, IL

```{r}
chi_bb <- c(-87.80, 41.75, -87.55, 42.00)
docks_raw_chi <- opq(bbox = chi_bb) |>
  add_osm_feature(key = "amenity", value = "bicycle_rental") |>
  add_osm_feature(key = "network", value = "Divvy") |>
  osmdata_sf()

docks_chi <- docks_raw_chi$osm_points |>
  st_as_sf() |>
  st_transform(3857) %>%
  filter(!is.na(network))
```

```{r}
hull_chi <- docks_chi |> 
  st_union() |> 
  st_concave_hull(.2)
hull_buffered_chi <- st_buffer(hull_chi, dist = 2500)
hull_buffered_wgs84_chi <- st_transform(hull_buffered_chi, 4326)
docks_chi <- st_intersection(
  st_transform(docks_chi, 4326),
  hull_buffered_wgs84_chi
)
```

```{r}
mapview(hull_buffered_wgs84_chi, col.regions = "lightblue", alpha.regions = 0.3) +
  mapview(docks_chi, col.regions = "red",
  layer.name = "Bikeshare Docks")
```
:::

The pattern of docks without any additional data already show us some differences in layout. New York and Chicago appear to have the most exhaustive bikeshare systems, with docks spanning nearly the entire geographic region with relative frequency everywhere. In Boston and Washington, docks seem to be more strategically placed with clusters extending away from the downtown area in various directions. Meanwhile, San Francisco has more sporadic placement with clustering only appearing in the downtown area.

While these maps all suggest divergent strategies for placing bikeshare docks, the next section will help pinpoint factors that may influence these decisions. Using Census data from `tidycensus` and train station locations from OpenStreetMap via the `osm` package, we can take a look at two variables that urban planners would likely be interested in, according to our literature review.

### Plotting Other Data

::: panel-tabset
### Washington, DC

```{r}
#| results: hide
#census_api_key("5e8593d6e289e6ae04c33a019f4951ca2bdd9523", install = TRUE)
dc_bg <- get_acs(
  geography = "block group",
  variables = "B01003_001E",
  year = 2022,
  survey = "acs5",
  state = c("DC", "MD", "VA"),
  geometry = TRUE
)
```

```{r}
dc_bg <- st_transform(dc_bg, st_crs(hull_buffered_wgs84_dc))

dc_bg <- dc_bg %>%
  mutate(area_original = as.numeric(st_area(geometry)))

dc_bg_clipped <- st_intersection(dc_bg, hull_buffered_wgs84_dc)

dc_bg_clipped <- dc_bg_clipped %>%
  mutate(
    area_clipped = as.numeric(st_area(geometry)),
    pop_adjusted = estimate * (area_clipped / area_original),
    area_km2 = area_clipped / 1e6,
    pop_density = pop_adjusted / area_km2
  )

dc_bg_density <- dc_bg_clipped %>%
  mutate(
    area_km2 = as.numeric(st_area(geometry)) / 1e6,
    pop_density = estimate / area_km2
  )
```

```{r}
max_density <- 20000
dc_bg_density$pop_density_plot <- pmin(dc_bg_density$pop_density, max_density)
```

```{r}
dc_bbox <- st_bbox(hull_buffered_wgs84_dc)

# bus_stops <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "bus_stop") %>%
#   osmdata_sf() %>%
#   .$osm_points %>%
#   st_transform(st_crs(hull_buffered_wgs84_dc))

stations_dc <- opq(dc_bbox) %>%
  add_osm_feature(key = "railway", value = "station") %>%
  osmdata_sf() %>%
  .$osm_points %>%
  st_transform(st_crs(hull_buffered_wgs84_dc))
```

```{r}
pal <- leaflet::colorNumeric(viridis(20), domain = dc_bg_density$pop_density_plot)
leaflet() %>%
  leaflet::addProviderTiles("CartoDB.Positron") %>%
  
  # Population density polygons
  leaflet::addPolygons(
    data = dc_bg_density,
    fillColor = ~pal(pop_density_plot),
    fillOpacity = 0.6,
    color = "grey",
    weight = 0.5,
    group = "Population Density",
    popup = ~paste0("Population density: ", round(pop_density_plot))
  ) %>%
  
  # Dock locations
  leaflet::addCircleMarkers(
    data = docks_dc,
    radius = 1,
    color = "blue",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Bikeshare Docks"
  ) %>%
  
  # Train stations
  leaflet::addCircleMarkers(
    data = stations_dc,
    radius = 1,
    color = "red",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Train Stations"
  ) %>%
  
  # Layer controls
  leaflet::addLayersControl(
    overlayGroups = c("Population Density", "Bikeshare Docks", "Train Stations"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
```

### New York, NY

```{r}
#| results: hide
nyc_bg <- get_acs(
  geography = "block group",
  variables = "B01003_001E",
  year = 2022,
  survey = "acs5",
  state = c("NY", "NJ"),
  geometry = TRUE
)
```

```{r}
nyc_bg <- st_transform(nyc_bg, st_crs(hull_buffered_wgs84_nyc))

nyc_bg <- nyc_bg %>%
  mutate(area_original = as.numeric(st_area(geometry)))

nyc_bg_clipped <- st_intersection(nyc_bg, hull_buffered_wgs84_nyc)

nyc_bg_clipped <- nyc_bg_clipped %>%
  mutate(
    area_clipped = as.numeric(st_area(geometry)),
    pop_adjusted = estimate * (area_clipped / area_original),
    area_km2 = area_clipped / 1e6,
    pop_density = pop_adjusted / area_km2
  )

nyc_bg_density <- nyc_bg_clipped %>%
  mutate(
    area_km2 = as.numeric(st_area(geometry)) / 1e6,
    pop_density = estimate / area_km2
  )
```

```{r}
max_density <- 75000
nyc_bg_density$pop_density_plot <- pmin(nyc_bg_density$pop_density, max_density)
```

```{r}
nyc_bbox <- st_bbox(hull_buffered_wgs84_nyc)

# bus_stops <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "bus_stop") %>%
#   osmdata_sf() %>%
#   .$osm_points %>%
#   st_transform(st_crs(hull_buffered_wgs84_dc))

stations_nyc <- opq(nyc_bbox) %>%
  add_osm_feature(key = "railway", value = "station") %>%
  osmdata_sf() %>%
  .$osm_points %>%
  st_transform(st_crs(hull_buffered_wgs84_nyc))
```

```{r}
pal <- leaflet::colorNumeric(viridis(20), domain = nyc_bg_density$pop_density_plot)
leaflet() %>%
  leaflet::addProviderTiles("CartoDB.Positron") %>%
  
  # Population density polygons
  leaflet::addPolygons(
    data = nyc_bg_density,
    fillColor = ~pal(pop_density_plot),
    fillOpacity = 0.6,
    color = "grey",
    weight = 0.5,
    group = "Population Density",
    popup = ~paste0("Population density: ", round(pop_density_plot))
  ) %>%
  
  # Dock locations
  leaflet::addCircleMarkers(
    data = docks_nyc,
    radius = 1,
    color = "blue",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Bikeshare Docks"
  ) %>%
  
  # Train stations
  leaflet::addCircleMarkers(
    data = stations_nyc,
    radius = 1,
    color = "red",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Train Stations"
  ) %>%
  
  # Layer controls
  leaflet::addLayersControl(
    overlayGroups = c("Population Density", "Bikeshare Docks", "Train Stations"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
```

### Boston, MA

```{r}
#| results: hide
bos_bg <- get_acs(
  geography = "block group",
  variables = "B01003_001E",
  year = 2022,
  survey = "acs5",
  state = c("MA"),
  geometry = TRUE
)
```

```{r}
bos_bg <- st_transform(bos_bg, st_crs(hull_buffered_wgs84_bos))

bos_bg <- bos_bg %>%
  mutate(area_original = as.numeric(st_area(geometry)))

bos_bg_clipped <- st_intersection(bos_bg, hull_buffered_wgs84_bos)

bos_bg_clipped <- bos_bg_clipped %>%
  mutate(
    area_clipped = as.numeric(st_area(geometry)),
    pop_adjusted = estimate * (area_clipped / area_original),
    area_km2 = area_clipped / 1e6,
    pop_density = pop_adjusted / area_km2
  )

bos_bg_density <- bos_bg_clipped %>%
  mutate(
    area_km2 = as.numeric(st_area(geometry)) / 1e6,
    pop_density = estimate / area_km2
  )
```

```{r}
max_density <- 25000
bos_bg_density$pop_density_plot <- pmin(bos_bg_density$pop_density, max_density)
```

```{r}
bos_bbox <- st_bbox(hull_buffered_wgs84_bos)

# bus_stops <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "bus_stop") %>%
#   osmdata_sf() %>%
#   .$osm_points %>%
#   st_transform(st_crs(hull_buffered_wgs84_dc))

stations_bos <- opq(bos_bbox) %>%
  add_osm_feature(key = "railway", value = "station") %>%
  osmdata_sf() %>%
  .$osm_points %>%
  st_transform(st_crs(hull_buffered_wgs84_bos))
```

```{r}
pal <- leaflet::colorNumeric(viridis(20), domain = bos_bg_density$pop_density_plot)
leaflet() %>%
  leaflet::addProviderTiles("CartoDB.Positron") %>%
  
  # Population density polygons
  leaflet::addPolygons(
    data = bos_bg_density,
    fillColor = ~pal(pop_density_plot),
    fillOpacity = 0.6,
    color = "grey",
    weight = 0.5,
    group = "Population Density",
    popup = ~paste0("Population density: ", round(pop_density_plot))
  ) %>%
  
  # Dock locations
  leaflet::addCircleMarkers(
    data = docks_bos,
    radius = 1,
    color = "blue",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Bikeshare Docks"
  ) %>%
  
  # Train stations
  leaflet::addCircleMarkers(
    data = stations_bos,
    radius = 1,
    color = "red",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Train Stations"
  ) %>%
  
  # Layer controls
  leaflet::addLayersControl(
    overlayGroups = c("Population Density", "Bikeshare Docks", "Train Stations"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
```

### San Francisco, CA

```{r}
#| results: hide
sf_bg <- get_acs(
  geography = "block group",
  variables = "B01003_001E",
  year = 2022,
  survey = "acs5",
  state = c("CA"),
  geometry = TRUE
)
```

```{r}
sf_bg <- st_transform(sf_bg, st_crs(hull_buffered_wgs84_sf))

sf_bg <- sf_bg %>%
  mutate(area_original = as.numeric(st_area(geometry)))

sf_bg_clipped <- st_intersection(sf_bg, hull_buffered_wgs84_sf)

sf_bg_clipped <- sf_bg_clipped %>%
  mutate(
    area_clipped = as.numeric(st_area(geometry)),
    pop_adjusted = estimate * (area_clipped / area_original),
    area_km2 = area_clipped / 1e6,
    pop_density = pop_adjusted / area_km2
  )

sf_bg_density <- sf_bg_clipped %>%
  mutate(
    area_km2 = as.numeric(st_area(geometry)) / 1e6,
    pop_density = estimate / area_km2
  )
```

```{r}
max_density <- 35000
sf_bg_density$pop_density_plot <- pmin(sf_bg_density$pop_density, max_density)
```

```{r}
sf_bbox <- st_bbox(hull_buffered_wgs84_sf)

# bus_stops <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "bus_stop") %>%
#   osmdata_sf() %>%
#   .$osm_points %>%
#   st_transform(st_crs(hull_buffered_wgs84_dc))

stations_sf <- opq(sf_bbox) %>%
  add_osm_feature(key = "railway", value = "station") %>%
  osmdata_sf() %>%
  .$osm_points %>%
  st_transform(st_crs(hull_buffered_wgs84_sf))
```

```{r}
pal <- leaflet::colorNumeric(viridis(20), domain = sf_bg_density$pop_density_plot)
leaflet() %>%
  leaflet::addProviderTiles("CartoDB.Positron") %>%
  
  # Population density polygons
  leaflet::addPolygons(
    data = sf_bg_density,
    fillColor = ~pal(pop_density_plot),
    fillOpacity = 0.6,
    color = "grey",
    weight = 0.5,
    group = "Population Density",
    popup = ~paste0("Population density: ", round(pop_density_plot))
  ) %>%
  
  # Dock locations
  leaflet::addCircleMarkers(
    data = docks_sf,
    radius = 1,
    color = "blue",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Bikeshare Docks"
  ) %>%
  
  # Train stations
  leaflet::addCircleMarkers(
    data = stations_sf,
    radius = 1,
    color = "red",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Train Stations"
  ) %>%
  
  # Layer controls
  leaflet::addLayersControl(
    overlayGroups = c("Population Density", "Bikeshare Docks", "Train Stations"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
```

### Chicago, IL

```{r}
#| results: hide
chi_bg <- get_acs(
  geography = "block group",
  variables = "B01003_001E",
  year = 2022,
  survey = "acs5",
  state = c("IL"),
  geometry = TRUE
)
```

```{r}
chi_bg <- st_transform(chi_bg, st_crs(hull_buffered_wgs84_chi))

chi_bg <- chi_bg %>%
  mutate(area_original = as.numeric(st_area(geometry)))

chi_bg_clipped <- st_intersection(chi_bg, hull_buffered_wgs84_chi)

chi_bg_clipped <- chi_bg_clipped %>%
  mutate(
    area_clipped = as.numeric(st_area(geometry)),
    pop_adjusted = estimate * (area_clipped / area_original),
    area_km2 = area_clipped / 1e6,
    pop_density = pop_adjusted / area_km2
  )

chi_bg_density <- chi_bg_clipped %>%
  mutate(
    area_km2 = as.numeric(st_area(geometry)) / 1e6,
    pop_density = estimate / area_km2
  )
```

```{r}
max_density <- 30000
chi_bg_density$pop_density_plot <- pmin(chi_bg_density$pop_density, max_density)
```

```{r}
chi_bbox <- st_bbox(hull_buffered_wgs84_chi)

# bus_stops <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "bus_stop") %>%
#   osmdata_sf() %>%
#   .$osm_points %>%
#   st_transform(st_crs(hull_buffered_wgs84_dc))

stations_chi <- opq(chi_bbox) %>%
  add_osm_feature(key = "railway", value = "station") %>%
  osmdata_sf() %>%
  .$osm_points %>%
  st_transform(st_crs(hull_buffered_wgs84_chi))
```

```{r}
pal <- leaflet::colorNumeric(viridis(20), domain = chi_bg_density$pop_density_plot)
leaflet() %>%
  leaflet::addProviderTiles("CartoDB.Positron") %>%
  
  # Population density polygons
  leaflet::addPolygons(
    data = chi_bg_density,
    fillColor = ~pal(pop_density_plot),
    fillOpacity = 0.6,
    color = "grey",
    weight = 0.5,
    group = "Population Density",
    popup = ~paste0("Population density: ", round(pop_density_plot))
  ) %>%
  
  # Dock locations
  leaflet::addCircleMarkers(
    data = docks_chi,
    radius = 1,
    color = "blue",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Bikeshare Docks"
  ) %>%
  
  # Train stations
  leaflet::addCircleMarkers(
    data = stations_chi,
    radius = 1,
    color = "red",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Train Stations"
  ) %>%
  
  # Layer controls
  leaflet::addLayersControl(
    overlayGroups = c("Population Density", "Bikeshare Docks", "Train Stations"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
```
:::

Adding train stations and population density to these maps starts provide some strategic insights as to where docks tend to cluster. Starting with Washington, D.C., the most heavily clustered areas reliably line up with dense populations that are often served by multiple train stations in close proximity. Meanwhile, the clusters of docks that stem out from the center of the city seem to line up nearly perfectly with train stations following the same pattern, indicating intentional design to follow existing public transit infrastructure.

As mentioned in the previous section, New York and Chicago have similarly exhaustive coverage over their city, both with a higher density of docks in central areas. However, Chicago's system seems to increase dock density exclusively in population-dense areas. New York, on the other hand, has its greatest dock density in lower Manhattan, even in neighborhoods like midtown which has low population density. In contrast with the relatively sparse coverage in high-population upper Manhattan neighborhoods, we can surmise that this system was partially designed with commercial neighborhoods taking precedent over residential ones. In both cases, public transit does not jump out as having an effect on dock locations.

Boston and San Francisco seem to diverge the most from population density, as Boston has a few dense neighborhoods that lack bikeshare coverage, and San Francisco's most dense area is curiously surrounded by bikeshare docks but doesn't contain them. Additionally, the docks seem to cover entirely different parts of the city than train stations in San Francisco, while no apparent pattern jumps out in Boston.

## Univariate Testing

Prior to hypothesis testing, another valuable form of exploratory data analysis can be to simply quantify differences in bikeshare dock deployment via clustering. This can be accomplished by calculating **Moran's I**, a statistic that measures spatial autocorrelation within a region to indicate if nodes are clustered, disperse, or somewhere in between. The following section will output Moran's I values for all five cities, along with a `leaflet` plot for each one that highlights areas of greater clustering than average.

::: panel-tabset
### Washington, DC

```{r}
coords_dc <- st_coordinates(docks_dc)

knn_dc <- knearneigh(coords_dc, k = 4)
nb_dc <- knn2nb(knn_dc)

lw_dc <- nb2listw(nb_dc, style = "W")

inv_dist_dc <- sapply(1:nrow(coords_dc), function(i) {
  neighbors_dc <- nb_dc[[i]]
  if(length(neighbors_dc) == 0) return(0)
  1 / mean(sqrt((coords_dc[i,1]-coords_dc[neighbors_dc,1])^2 + (coords_dc[i,2]-coords_dc[neighbors_dc,2])^2))
})

moran_result_dc <- moran.test(inv_dist_dc, lw_dc, zero.policy = TRUE)
moran_result_dc
```

```{r}
dock_values_dc <- inv_dist_dc

localI_dc <- localmoran(dock_values_dc, lw_dc, zero.policy = TRUE)
docks_dc$localI <- localI_dc[,1]

leaflet(docks_dc) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    radius = 4,
    color = ~viridis::viridis(20)[cut(localI, breaks = 20)],
    fillOpacity = 0.8,
    popup = ~paste0("Local Moran's I: ", round(localI, 3))
  )
```

### New York, NY

```{r}
coords_nyc <- st_coordinates(docks_nyc)

knn_nyc <- knearneigh(coords_nyc, k = 4)
nb_nyc <- knn2nb(knn_nyc)

lw_nyc <- nb2listw(nb_nyc, style = "W")

inv_dist_nyc <- sapply(1:nrow(coords_nyc), function(i) {
  neighbors_nyc <- nb_nyc[[i]]
  if(length(neighbors_nyc) == 0) return(0)
  1 / mean(sqrt((coords_nyc[i,1]-coords_nyc[neighbors_nyc,1])^2 + (coords_nyc[i,2]-coords_nyc[neighbors_nyc,2])^2))
})

moran_result_nyc <- moran.test(inv_dist_nyc, lw_nyc, zero.policy = TRUE)
moran_result_nyc
```

```{r}
dock_values_nyc <- inv_dist_nyc

localI_nyc <- localmoran(dock_values_nyc, lw_nyc, zero.policy = TRUE)
docks_nyc$localI <- localI_nyc[,1]

leaflet(docks_nyc) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    radius = 4,
    color = ~viridis::viridis(20)[cut(localI, breaks = 20)],
    fillOpacity = 0.8,
    popup = ~paste0("Local Moran's I: ", round(localI, 3))
  )
```

### Boston, MA

```{r}
coords_bos <- st_coordinates(docks_bos)

knn_bos <- knearneigh(coords_bos, k = 4)
nb_bos <- knn2nb(knn_bos)

lw_bos <- nb2listw(nb_bos, style = "W")

inv_dist_bos <- sapply(1:nrow(coords_bos), function(i) {
  neighbors_bos <- nb_bos[[i]]
  if(length(neighbors_bos) == 0) return(0)
  1 / mean(sqrt((coords_bos[i,1]-coords_bos[neighbors_bos,1])^2 + (coords_bos[i,2]-coords_bos[neighbors_bos,2])^2))
})

moran_result_bos <- moran.test(inv_dist_bos, lw_bos, zero.policy = TRUE)
moran_result_bos
```

```{r}
dock_values_bos <- inv_dist_bos

localI_bos <- localmoran(dock_values_bos, lw_bos, zero.policy = TRUE)
docks_bos$localI <- localI_bos[,1]

leaflet(docks_bos) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    radius = 4,
    color = ~viridis::viridis(20)[cut(localI, breaks = 20)],
    fillOpacity = 0.8,
    popup = ~paste0("Local Moran's I: ", round(localI, 3))
  )
```

### San Francisco, CA

```{r}
coords_sf <- st_coordinates(docks_sf)

knn_sf <- knearneigh(coords_sf, k = 4)
nb_sf <- knn2nb(knn_sf)

lw_sf <- nb2listw(nb_sf, style = "W")

inv_dist_sf <- sapply(1:nrow(coords_sf), function(i) {
  neighbors_sf <- nb_sf[[i]]
  if(length(neighbors_sf) == 0) return(0)
  1 / mean(sqrt((coords_sf[i,1]-coords_sf[neighbors_sf,1])^2 + (coords_sf[i,2]-coords_sf[neighbors_sf,2])^2))
})

moran_result_sf <- moran.test(inv_dist_sf, lw_sf, zero.policy = TRUE)
moran_result_sf
```

```{r}
dock_values_sf <- inv_dist_sf

localI_sf <- localmoran(dock_values_sf, lw_sf, zero.policy = TRUE)
docks_sf$localI <- localI_sf[,1]

leaflet(docks_sf) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    radius = 4,
    color = ~viridis::viridis(20)[cut(localI, breaks = 20)],
    fillOpacity = 0.8,
    popup = ~paste0("Local Moran's I: ", round(localI, 3))
  )
```

### Chicago, IL

```{r}
coords_chi <- st_coordinates(docks_chi)

knn_chi <- knearneigh(coords_chi, k = 4)
nb_chi <- knn2nb(knn_chi)

lw_chi <- nb2listw(nb_chi, style = "W")

inv_dist_chi <- sapply(1:nrow(coords_chi), function(i) {
  neighbors_chi <- nb_chi[[i]]
  if(length(neighbors_chi) == 0) return(0)
  1 / mean(sqrt((coords_chi[i,1]-coords_chi[neighbors_chi,1])^2 + (coords_chi[i,2]-coords_chi[neighbors_chi,2])^2))
})

moran_result_chi <- moran.test(inv_dist_chi, lw_chi, zero.policy = TRUE)
moran_result_chi
```

```{r}
dock_values_chi <- inv_dist_chi

localI_chi <- localmoran(dock_values_chi, lw_chi, zero.policy = TRUE)
docks_chi$localI <- localI_chi[,1]

leaflet(docks_chi) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    radius = 4,
    color = ~viridis::viridis(20)[cut(localI, breaks = 20)],
    fillOpacity = 0.8,
    popup = ~paste0("Local Moran's I: ", round(localI, 3))
  )
```
:::

### Univariate Results

| Metropolitan Area | Moran's I Statistic |
|-------------------|---------------------|
| Washington, DC    | $0.8418107214$      |
| New York, NY      | $0.6827748494$      |
| Boston, MA        | $0.842640558$       |
| San Francisco, CA | $0.808442476$       |
| Chicago, IL       | $0.9584339511$      |

These maps and statistics mostly confirm some of the observations from the EDA section. The results to note are the differences outlined between New York and Chicago. Chicago's system, while spread out over the city, clearly is characterized by intense clustering in the downtown, such that it produces the highest Moran's I of the bunch. Meanwhile, New York's Moran's I tells us that most of the docks are intentionally spaced out more than in other cities, providing coverage to more neighborhoods and less favoritism towards certain regions.

The other observation from this table is that all of the Moran's I statistics are relatively high, confirming a common-sense assumption that bikeshare docks tend to be placed in the presence of other bikeshare docks.

## Hypothesis Test: Population Density

Our first hypothesis tests will pertain to population density. As a reminder the null and alternative hypotheses are:

-   **Null** **hypothesis**: There is no relationship between the spatial distribution of bikeshare docks and population density. Dock locations occur independently of areas with higher or lower population density.
-   **Alternative** **hypothesis**: There is a relationship between the spatial distribution of Capital Bikeshare docks and population density. Dock locations are more likely to occur in areas with higher population density.

To test these, we will first look at correlation functions to quantify the relationship between dock locations and population density. Specifically, we will be using a likelihood ratio test within the framework of a Poisson Point Process to get both the statistical significance of the population density coefficient and a likelihood ratio that compares a model that relies on population density to a null model that is spatially homogeneous. The following section contains the process for executing these tests, with a table displaying relevant results at the end.

### Correlation Functions

::: panel-tabset
### Washington, DC

```{r}
proj_crs <- "EPSG:32618"
docks_proj_dc <- st_transform(docks_dc, 32618)
boundary_proj_dc <- st_transform(hull_buffered_wgs84_dc, 32618)
boundary_owin_dc <- as.owin(boundary_proj_dc)

dock_coords_dc <- st_coordinates(docks_proj_dc)

dock_ppp_dc <- ppp(
  x = dock_coords_dc[,1],
  y = dock_coords_dc[,2],
  window = boundary_owin_dc
)
dc_bg_density_proj <- st_transform(dc_bg_density, proj_crs)
rtemplate_dc <- rast(
  extent = ext(dc_bg_density_proj),
  resolution = 250,
  crs = "EPSG:32618"
)

pop_rast_dc <- rasterize(
  vect(dc_bg_density_proj),
  rtemplate_dc,
  field = "pop_density_plot",
  fun = "mean"
)

pop_rast_dc <- mask(pop_rast_dc, vect(boundary_proj_dc))
```

```{r}
nx_dc <- ncol(pop_rast_dc)
ny_dc <- nrow(pop_rast_dc)
r_ext_dc <- ext(pop_rast_dc)
res_x_dc <- res(pop_rast_dc)[1]
res_y_dc <- res(pop_rast_dc)[2]

xcol_dc <- seq(r_ext_dc$xmin + res_x_dc/2, r_ext_dc$xmax - res_x_dc/2, length.out = nx_dc)
yrow_dc <- seq(r_ext_dc$ymin + res_y_dc/2, r_ext_dc$ymax - res_y_dc/2, length.out = ny_dc)

vals_dc <- values(pop_rast_dc, mat = FALSE)
mat_dc <- matrix(vals_dc, nrow = ny_dc, ncol = nx_dc, byrow = TRUE)

mat_oriented_dc <- mat_dc[ny_dc:1, , drop = FALSE]
pop_im_dc <- im(mat_oriented_dc, xcol_dc, yrow_dc)
```

```{r}
m0_dc <- ppm(dock_ppp_dc ~ 1)
m1_dc <- ppm(dock_ppp_dc ~ pop_im_dc)
summary(m1_dc)
anova(m0_dc, m1_dc, test="LR")
```

### New York, NY

```{r}
docks_proj_nyc <- st_transform(docks_nyc, 32618)
boundary_proj_nyc <- st_transform(hull_buffered_wgs84_nyc, 32618)
boundary_owin_nyc <- as.owin(boundary_proj_nyc)

dock_coords_nyc <- st_coordinates(docks_proj_nyc)

dock_ppp_nyc <- ppp(
  x = dock_coords_nyc[,1],
  y = dock_coords_nyc[,2],
  window = boundary_owin_nyc
)
nyc_bg_density_proj <- st_transform(nyc_bg_density, crs = proj_crs)

rtemplate_nyc <- rast(
  extent = ext(nyc_bg_density_proj),
  resolution = 250,
  crs = "EPSG:32618"
)

pop_rast_nyc <- rasterize(
  vect(nyc_bg_density_proj),
  rtemplate_nyc,
  field = "pop_density_plot",
  fun = "mean"
)
pop_rast_nyc <- mask(pop_rast_nyc, vect(boundary_proj_nyc))
```

```{r}
nx_nyc <- ncol(pop_rast_nyc)
ny_nyc <- nrow(pop_rast_nyc)
r_ext_nyc <- ext(pop_rast_nyc)
res_x_nyc <- res(pop_rast_nyc)[1]
res_y_nyc <- res(pop_rast_nyc)[2]

xcol_nyc <- seq(r_ext_nyc$xmin + res_x_nyc/2, r_ext_nyc$xmax - res_x_nyc/2, length.out = nx_nyc)
yrow_nyc <- seq(r_ext_nyc$ymin + res_y_nyc/2, r_ext_nyc$ymax - res_y_nyc/2, length.out = ny_nyc)

vals_nyc <- values(pop_rast_nyc, mat = FALSE)
mat_nyc <- matrix(vals_nyc, nrow = ny_nyc, ncol = nx_nyc, byrow = TRUE)

mat_oriented_nyc <- mat_nyc[ny_nyc:1, , drop = FALSE]
pop_im_nyc <- im(mat_oriented_nyc, xcol_nyc, yrow_nyc)
```

```{r}
m0_nyc <- ppm(dock_ppp_nyc ~ 1)
m1_nyc <- ppm(dock_ppp_nyc ~ pop_im_nyc)
summary(m1_nyc)
anova(m0_nyc, m1_nyc, test="LR")
```

### Boston, MA

```{r}
docks_proj_bos <- st_transform(docks_bos, 32618)
boundary_proj_bos <- st_transform(hull_buffered_wgs84_bos, 32618)
boundary_owin_bos <- as.owin(boundary_proj_bos)

dock_coords_bos <- st_coordinates(docks_proj_bos)

dock_ppp_bos <- ppp(
  x = dock_coords_bos[,1],
  y = dock_coords_bos[,2],
  window = boundary_owin_bos
)
bos_bg_density_proj <- st_transform(bos_bg_density, crs = proj_crs)

rtemplate_bos <- rast(
  extent = ext(bos_bg_density_proj),
  resolution = 250,
  crs = "EPSG:32618"
)

pop_rast_bos <- rasterize(
  vect(bos_bg_density_proj),
  rtemplate_bos,
  field = "pop_density_plot",
  fun = "mean"
)
pop_rast_bos <- mask(pop_rast_bos, vect(boundary_proj_bos))
```

```{r}
nx_bos <- ncol(pop_rast_bos)
ny_bos <- nrow(pop_rast_bos)
r_ext_bos <- ext(pop_rast_bos)
res_x_bos <- res(pop_rast_bos)[1]
res_y_bos <- res(pop_rast_bos)[2]

xcol_bos <- seq(r_ext_bos$xmin + res_x_bos/2, r_ext_bos$xmax - res_x_bos/2, length.out = nx_bos)
yrow_bos <- seq(r_ext_bos$ymin + res_y_bos/2, r_ext_bos$ymax - res_y_bos/2, length.out = ny_bos)

vals_bos <- values(pop_rast_bos, mat = FALSE)
mat_bos <- matrix(vals_bos, nrow = ny_bos, ncol = nx_bos, byrow = TRUE)

mat_oriented_bos <- mat_bos[ny_bos:1, , drop = FALSE]
pop_im_bos <- im(mat_oriented_bos, xcol_bos, yrow_bos)
```

```{r}
m0_bos <- ppm(dock_ppp_bos ~ 1)
m1_bos <- ppm(dock_ppp_bos ~ pop_im_bos)
summary(m1_bos)
anova(m0_bos, m1_bos, test="LR")
```

### San Francisco, CA

```{r}
docks_proj_sf <- st_transform(docks_sf, 32618)
boundary_proj_sf <- st_transform(hull_buffered_wgs84_sf, 32618)
boundary_owin_sf <- as.owin(boundary_proj_sf)

dock_coords_sf <- st_coordinates(docks_proj_sf)

dock_ppp_sf <- ppp(
  x = dock_coords_sf[,1],
  y = dock_coords_sf[,2],
  window = boundary_owin_sf
)
sf_bg_density_proj <- st_transform(sf_bg_density, crs = proj_crs)

rtemplate_sf <- rast(
  extent = ext(sf_bg_density_proj),
  resolution = 250,
  crs = "EPSG:32618"
)

pop_rast_sf <- rasterize(
  vect(sf_bg_density_proj),
  rtemplate_sf,
  field = "pop_density_plot",
  fun = "mean"
)
pop_rast_sf <- mask(pop_rast_sf, vect(boundary_proj_sf))
```

```{r}
nx_sf <- ncol(pop_rast_sf)
ny_sf <- nrow(pop_rast_sf)
r_ext_sf <- ext(pop_rast_sf)
res_x_sf <- res(pop_rast_sf)[1]
res_y_sf <- res(pop_rast_sf)[2]

xcol_sf <- seq(r_ext_sf$xmin + res_x_sf/2, r_ext_sf$xmax - res_x_sf/2, length.out = nx_sf)
yrow_sf <- seq(r_ext_sf$ymin + res_y_sf/2, r_ext_sf$ymax - res_y_sf/2, length.out = ny_sf)

vals_sf <- values(pop_rast_sf, mat = FALSE)
mat_sf <- matrix(vals_sf, nrow = ny_sf, ncol = nx_sf, byrow = TRUE)

mat_oriented_sf <- mat_sf[ny_sf:1, , drop = FALSE]
pop_im_sf <- im(mat_oriented_sf, xcol_sf, yrow_sf)
```

```{r}
m0_sf <- ppm(dock_ppp_sf ~ 1)
m1_sf <- ppm(dock_ppp_sf ~ pop_im_sf)
summary(m1_sf)
anova(m0_sf, m1_sf, test="LR")
```

### Chicago, IL

```{r}
docks_proj_chi <- st_transform(docks_chi, 32618)
boundary_proj_chi <- st_transform(hull_buffered_wgs84_chi, 32618)
boundary_owin_chi <- as.owin(boundary_proj_chi)

dock_coords_chi <- st_coordinates(docks_proj_chi)

dock_ppp_chi <- ppp(
  x = dock_coords_chi[,1],
  y = dock_coords_chi[,2],
  window = boundary_owin_chi
)
chi_bg_density_proj <- st_transform(chi_bg_density, crs = proj_crs)

rtemplate_chi <- rast(
  extent = ext(chi_bg_density_proj),
  resolution = 250,
  crs = "EPSG:32618"
)

pop_rast_chi <- rasterize(
  vect(chi_bg_density_proj),
  rtemplate_chi,
  field = "pop_density_plot",
  fun = "mean"
)
pop_rast_chi <- mask(pop_rast_chi, vect(boundary_proj_chi))
```

```{r}
nx_chi <- ncol(pop_rast_chi)
ny_chi <- nrow(pop_rast_chi)
r_ext_chi <- ext(pop_rast_chi)
res_x_chi <- res(pop_rast_chi)[1]
res_y_chi <- res(pop_rast_chi)[2]

xcol_chi <- seq(r_ext_chi$xmin + res_x_chi/2, r_ext_chi$xmax - res_x_chi/2, length.out = nx_chi)
yrow_chi <- seq(r_ext_chi$ymin + res_y_chi/2, r_ext_chi$ymax - res_y_chi/2, length.out = ny_chi)

vals_chi <- values(pop_rast_chi, mat = FALSE)
mat_chi <- matrix(vals_chi, nrow = ny_chi, ncol = nx_chi, byrow = TRUE)

mat_oriented_chi <- mat_chi[ny_chi:1, , drop = FALSE]
pop_im_chi <- im(mat_oriented_chi, xcol_chi, yrow_chi)
```

```{r}
m0_chi <- ppm(dock_ppp_chi ~ 1)
m1_chi <- ppm(dock_ppp_chi ~ pop_im_chi)
summary(m1_chi)
anova(m0_chi, m1_chi, test="LR")
```
:::

### Correlation Function Results against Population Density

| Metropolitan Area | Docks   | `pop_im` coefficient   | Z-value | Deviance diff |
|-------------------|---------|------------------------|---------|---------------|
| Washington, DC    | $780$   | $1.530 \times 10^{-4}$ | $25.07$ | $423.41$      |
| New York, NY      | $2,303$ | $1.590 \times 10^{-5}$ | $16.16$ | $235.86$      |
| Boston, MA        | $320$   | $5.221 \times 10^{-5}$ | $6.32$  | $35.81$       |
| San Francisco, CA | $107$   | $3.529 \times 10^{-5}$ | $2.76$  | $6.96$        |
| Chicago, IL       | $909$   | $8.919 \times 10^{-5}$ | $18.61$ | $248.94$      |

This table offers us a good summation of how bikeshare dock locations correlate to population density. The first thing to note is that each test placed great importance on the population density coefficient, confirming that population density is positively correlated with docks in all five cities. However, the specific Z-values tell us that the extent of these correlations are not the same.

Going in descending order of Z-value, Washington's system most closely follows population density, followed by Chicago and New York, while Boston and Chicago's docks are relatively loosely tied to where people live.

### Monte Carlo Simulations

In addition to our Poisson Point Process, performing Monte Carlo simulations can provide extra confirmation of our findings by comparing the real-world deployments of bikeshare docks with randomized placements based on controlled conditions. For testing this hypothesis, all we need to do is simulate randomized placements that are based on no outside bias and see if we get results that are more or less tethered to population density. This will be quantified by checking how the population density coefficients compare to the real value. The proportion of simulated coefficients that exceed the real coefficient will serve as our p-value. The following section executes 999 simulations for each city, with a table of results at the end.

::: panel-tabset
### Washington, DC

```{r}
set.seed(202)
model_obs_dc <- ppm(dock_ppp_dc ~ pop_im_dc)
coef_obs_dc <- coef(model_obs_dc)["pop_im_dc"]

nsim <- 999
coef_sim_dc <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_dc <- runifpoint(n = dock_ppp_dc$n, win = boundary_owin_dc)
  sim_model_dc <- ppm(sim_ppp_dc ~ pop_im_dc)
  coef_sim_dc[i] <- coef(sim_model_dc)["pop_im_dc"]
}

p_value_dc <- (sum(coef_sim_dc >= coef_obs_dc) + 1) / (nsim + 1)

hist(
  coef_sim_dc,
  main = "Washington, DC: Monte Carlo Distribution of Population Density Coefficients",
  xlab = "Simulated Population Density Coefficient",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(min(coef_sim_dc, coef_obs_dc), max(coef_sim_dc, coef_obs_dc))
)
abline(v = coef_obs_dc, col = "red", lwd = 3)

p_value_dc <- (sum(coef_sim_dc >= coef_obs_dc) + 1) / (nsim + 1)
p_value_dc
```

### New York, NY

```{r}
model_obs_nyc <- ppm(dock_ppp_nyc ~ pop_im_nyc)
coef_obs_nyc <- coef(model_obs_nyc)["pop_im_nyc"]

nsim <- 999
coef_sim_nyc <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_nyc <- runifpoint(n = dock_ppp_nyc$n, win = boundary_owin_nyc)
  sim_model_nyc <- ppm(sim_ppp_nyc ~ pop_im_nyc)
  coef_sim_nyc[i] <- coef(sim_model_nyc)["pop_im_nyc"]
}

p_value_nyc <- (sum(coef_sim_nyc >= coef_obs_nyc) + 1) / (nsim + 1)

hist(
  coef_sim_nyc,
  main = "New York, NY: Monte Carlo Distribution of Population Density Coefficients",
  xlab = "Simulated Population Density Coefficient",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(min(coef_sim_nyc, coef_obs_nyc), max(coef_sim_nyc, coef_obs_nyc))
)
abline(v = coef_obs_nyc, col = "red", lwd = 3)

p_value_nyc <- (sum(coef_sim_nyc >= coef_obs_nyc) + 1) / (nsim + 1)
p_value_nyc
```

### Boston, MA

```{r}
model_obs_bos <- ppm(dock_ppp_bos ~ pop_im_bos)
coef_obs_bos <- coef(model_obs_bos)["pop_im_bos"]

nsim <- 999
coef_sim_bos <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_bos <- runifpoint(n = dock_ppp_bos$n, win = boundary_owin_bos)
  sim_model_bos <- ppm(sim_ppp_bos ~ pop_im_bos)
  coef_sim_bos[i] <- coef(sim_model_bos)["pop_im_bos"]
}

p_value_bos <- (sum(coef_sim_bos >= coef_obs_bos) + 1) / (nsim + 1)

hist(
  coef_sim_bos,
  main = "Boston, MA: Monte Carlo Distribution of Population Density Coefficients",
  xlab = "Simulated Population Density Coefficient",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(min(coef_sim_bos, coef_obs_bos), max(coef_sim_bos, coef_obs_bos))
)
abline(v = coef_obs_bos, col = "red", lwd = 3)

p_value_bos <- (sum(coef_sim_bos >= coef_obs_bos) + 1) / (nsim + 1)
p_value_bos
```

### San Francisco, CA

```{r}
model_obs_sf <- ppm(dock_ppp_sf ~ pop_im_sf)
coef_obs_sf <- coef(model_obs_sf)["pop_im_sf"]

nsim <- 999
coef_sim_sf <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_sf <- runifpoint(n = dock_ppp_sf$n, win = boundary_owin_sf)
  sim_model_sf <- ppm(sim_ppp_sf ~ pop_im_sf)
  coef_sim_sf[i] <- coef(sim_model_sf)["pop_im_sf"]
}

p_value_sf <- (sum(coef_sim_sf >= coef_obs_sf) + 1) / (nsim + 1)

hist(
  coef_sim_sf,
  main = "San Francisco, CA: Monte Carlo Distribution of Population Density Coefficients",
  xlab = "Simulated Population Density Coefficient",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(min(coef_sim_sf, coef_obs_sf), max(coef_sim_sf, coef_obs_sf))
)
abline(v = coef_obs_sf, col = "red", lwd = 3)

p_value_sf <- (sum(coef_sim_sf >= coef_obs_sf) + 1) / (nsim + 1)
p_value_sf
```

### Chicago, IL

```{r}
model_obs_chi <- ppm(dock_ppp_chi ~ pop_im_chi)
coef_obs_chi <- coef(model_obs_chi)["pop_im_chi"]

nsim <- 999
coef_sim_chi <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_chi <- runifpoint(n = dock_ppp_chi$n, win = boundary_owin_chi)
  sim_model_chi <- ppm(sim_ppp_chi ~ pop_im_chi)
  coef_sim_chi[i] <- coef(sim_model_chi)["pop_im_chi"]
}

p_value_chi <- (sum(coef_sim_chi >= coef_obs_chi) + 1) / (nsim + 1)

hist(
  coef_sim_chi,
  main = "Chicago, IL: Monte Carlo Distribution of Population Density Coefficients",
  xlab = "Simulated Population Density Coefficient",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(min(coef_sim_chi, coef_obs_chi), max(coef_sim_chi, coef_obs_chi))
)
abline(v = coef_obs_chi, col = "red", lwd = 3)

p_value_chi <- (sum(coef_sim_chi >= coef_obs_chi) + 1) / (nsim + 1)
p_value_chi
```
:::

### Monte Carlo Simulation Results against Population Density

| Metropolitan Area | p-value |
|-------------------|---------|
| Washington, DC    | $0.001$ |
| New York, NY      | $0.001$ |
| Boston, MA        | $0.001$ |
| San Francisco, CA | $0.011$ |
| Chicago, IL       | $0.001$ |

This process, while it emphatically confirms our alternate hypothesis in almost all cases, proves less useful than deriving statistics via a Poisson Point Process. Looking at the histogram of population density coefficients, the real coefficient is comically far from all simulated results in Washington, New York, and Chicago. The interesting finding here is that San Francisco did produce some simulations with coefficients that exceeded the real value. This once again confirms that San Francisco's bikeshare system is uniquely only loosely tied to population density.

## Hypothesis Test: Train Stations

While testing bikeshare dock locations against population density revealed some interesting differences in deployment strategy, it is reasonable to presume that there are other factors that influence these decisions. As stated in the introduction and literature review, considering access to public transit can tell us a lot about the purpose of bikeshare: whether it targets people who lack other car-free transit options, serves as a complement to public transit, or does not consider this factor at all. Thus, we will be testing the following hypotheses:

-   **Null hypothesis**: There is no relationship between the spatial distribution of bikeshare docks and the presence of other public transit nodes.

-   **Alternative hypothesis:** There is a relationship between the spatial distribution of bikeshare docks and the presence of other public transit nodes. Dock locations are more likely to be placed in areas with existing public transit infrastructure.

The first method of testing these hypotheses will be similar to the correlation functions from the previous section. We will again use Poisson Point Processes to obtain the coefficient of the distance from each dock to its nearest train station, as well as the likelihood ratio to compare to a null model.

### Correlation Functions

::: panel-tabset
### Washington, DC

```{r}
stations_proj_dc <- st_transform(stations_dc, 32618)
docks_proj_dc <- st_transform(docks_dc, 32618)
boundary_proj_dc <- st_transform(hull_buffered_wgs84_dc, 32618)

boundary_owin_dc <- as.owin(boundary_proj_dc)
dock_coords_dc <- st_coordinates(docks_proj_dc)
dock_ppp_dc <- ppp(x = dock_coords_dc[,1],
                y = dock_coords_dc[,2],
                window = boundary_owin_dc)

station_coords_dc <- st_coordinates(stations_proj_dc)
station_ppp_dc <- ppp(x = station_coords_dc[,1],
                   y = station_coords_dc[,2],
                   window = boundary_owin_dc)

dist_im_dc <- distmap(station_ppp_dc)
```

```{r}
m0_dc <- ppm(dock_ppp_dc ~ 1)
m1_dc <- ppm(dock_ppp_dc ~ dist_im_dc)

anova(m0_dc, m1_dc, test = "LR")
summary(m1_dc)
```

### New York, NY

```{r}
stations_proj_nyc <- st_transform(stations_nyc, 32618)
docks_proj_nyc <- st_transform(docks_nyc, 32618)
boundary_proj_nyc <- st_transform(hull_buffered_wgs84_nyc, 32618)

boundary_owin_nyc <- as.owin(boundary_proj_nyc)
dock_coords_nyc <- st_coordinates(docks_proj_nyc)
dock_ppp_nyc <- ppp(x = dock_coords_nyc[,1],
                y = dock_coords_nyc[,2],
                window = boundary_owin_nyc)

station_coords_nyc <- st_coordinates(stations_proj_nyc)
station_ppp_nyc <- ppp(x = station_coords_nyc[,1],
                   y = station_coords_nyc[,2],
                   window = boundary_owin_nyc)

dist_im_nyc <- distmap(station_ppp_nyc)
```

```{r}
m0_nyc <- ppm(dock_ppp_nyc ~ 1)
m1_nyc <- ppm(dock_ppp_nyc ~ dist_im_nyc)

anova(m0_nyc, m1_nyc, test = "LR")
summary(m1_nyc)
```

### Boston, MA

```{r}
stations_proj_bos <- st_transform(stations_bos, 32618)
docks_proj_bos <- st_transform(docks_bos, 32618)
boundary_proj_bos <- st_transform(hull_buffered_wgs84_bos, 32618)

boundary_owin_bos <- as.owin(boundary_proj_bos)
dock_coords_bos <- st_coordinates(docks_proj_bos)
dock_ppp_bos <- ppp(x = dock_coords_bos[,1],
                y = dock_coords_bos[,2],
                window = boundary_owin_bos)

station_coords_bos <- st_coordinates(stations_proj_bos)
station_ppp_bos <- ppp(x = station_coords_bos[,1],
                   y = station_coords_bos[,2],
                   window = boundary_owin_bos)

dist_im_bos <- distmap(station_ppp_bos)
```

```{r}
m0_bos <- ppm(dock_ppp_bos ~ 1)
m1_bos <- ppm(dock_ppp_bos ~ dist_im_bos)

anova(m0_bos, m1_bos, test = "LR")
summary(m1_bos)
```

### San Francisco, CA

```{r}
stations_proj_sf <- st_transform(stations_sf, 32618)
docks_proj_sf <- st_transform(docks_sf, 32618)
boundary_proj_sf <- st_transform(hull_buffered_wgs84_sf, 32618)

boundary_owin_sf <- as.owin(boundary_proj_sf)
dock_coords_sf <- st_coordinates(docks_proj_sf)
dock_ppp_sf <- ppp(x = dock_coords_sf[,1],
                y = dock_coords_sf[,2],
                window = boundary_owin_sf)

station_coords_sf <- st_coordinates(stations_proj_sf)
station_ppp_sf <- ppp(x = station_coords_sf[,1],
                   y = station_coords_sf[,2],
                   window = boundary_owin_sf)

dist_im_sf <- distmap(station_ppp_sf)
```

```{r}
m0_sf <- ppm(dock_ppp_sf ~ 1)
m1_sf <- ppm(dock_ppp_sf ~ dist_im_sf)

anova(m0_sf, m1_sf, test = "LR")
summary(m1_sf)
```

### Chicago, IL

```{r}
stations_proj_chi <- st_transform(stations_chi, 32618)
docks_proj_chi <- st_transform(docks_chi, 32618)
boundary_proj_chi <- st_transform(hull_buffered_wgs84_chi, 32618)

boundary_owin_chi <- as.owin(boundary_proj_chi)
dock_coords_chi <- st_coordinates(docks_proj_chi)
dock_ppp_chi <- ppp(x = dock_coords_chi[,1],
                y = dock_coords_chi[,2],
                window = boundary_owin_chi)

station_coords_chi <- st_coordinates(stations_proj_chi)
station_ppp_chi <- ppp(x = station_coords_chi[,1],
                   y = station_coords_chi[,2],
                   window = boundary_owin_chi)

dist_im_chi <- distmap(station_ppp_chi)
```

```{r}
m0_chi <- ppm(dock_ppp_chi ~ 1)
m1_chi <- ppm(dock_ppp_chi ~ dist_im_chi)

anova(m0_chi, m1_chi, test = "LR")
summary(m1_chi)
```
:::

### Correlation Function Results against Train Station Locations

| Metropolitan Area | Docks   | `dist_im` coefficient   | Z-value  | Deviance diff |
|---------------|---------------|---------------|---------------|---------------|
| Washington, DC    | $774$   | $-9.08 \times 10^{-4}$  | $-19.30$ | $584.99$      |
| New York, NY      | $2,303$ | $-1.22 \times 10^{-3}$  | $-18.27$ | $420.63$      |
| Boston, MA        | $320$   | $-9.88 \times 10^{-4}$  | $-8.88$  | $115.03$      |
| San Francisco, CA | $107$   | $-2.91 \times 10^{-04}$ | $-2.76$  | $9.337$       |
| Chicago, IL       | $909$   | $-7.70 \times 10^{-4}$  | $-15.53$ | $305.55$      |

As we saw with the population density tests, all five null hypotheses are rejected by this process. The extent of how strong the coefficient representing distance to the nearest train station varies in a similar way to the population density coefficient, with the same cities experiencing roughly the same correlation. If we assume that population density and train station locations are heavily correlated, these results are heavily expected, so simply performing this test was not sufficient in isolating the presence of public transit.

### Monte Carlo Simulations

Similar to our other hypothesis test, running Monte Carlo simulations will help us better understand how real data compares to data that we can generate under specified conditions. Unlike the previous section, we will use this to distinguish between systems that happen to place docks close to train stations by nature of following population density from those that intentionally seek out locations with more or less access to public transit.

Therefore, in these 999 Monte Carlo simulations for each city, we will generate simulations that place docks proportionally to population density. We can then compare the coefficients for distance to the nearest train station, seeing if the real data has a stronger dependence on this variable than data with a bias towards denser areas. The process and results of these tests are below.

::: panel-tabset
### Washington, DC

```{r}
stations_proj_dc <- st_transform(stations_dc, 32618)
station_coords_dc <- st_coordinates(stations_proj_dc)

station_ppp_dc <- ppp(
  x = station_coords_dc[,1],
  y = station_coords_dc[,2],
  window = boundary_owin_dc
)

obs_dist_dc <- nndist(dock_ppp_dc, station_ppp_dc)
obs_mean_dist_dc <- mean(obs_dist_dc)


nsim <- 999
sim_mean_dist_dc <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_dc <- rpoint(
    n = dock_ppp_dc$n,
    f = pop_im_dc,
    win = boundary_owin_dc
  )
  sim_dist <- nndist(sim_ppp_dc, station_ppp_dc)
  sim_mean_dist_dc[i] <- mean(sim_dist)
}


p_value_dc <- (sum(sim_mean_dist_dc <= obs_mean_dist_dc) + 1) / (nsim + 1)

x_min <- min(sim_mean_dist_dc, obs_mean_dist_dc)
x_max <- max(sim_mean_dist_dc, obs_mean_dist_dc)

hist(
  sim_mean_dist_dc,
  main = "Washington, DC: Monte Carlo Distribution of\nDistance from Docks to Nearest Train Station",
  xlab = "Simulated Mean Distance",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(x_min, x_max)
)

abline(v = obs_mean_dist_dc, col = "red", lwd = 3)
p_value_dc
```

### New York, NY

```{r}
stations_proj_nyc <- st_transform(stations_nyc, 32618)
station_coords_nyc <- st_coordinates(stations_proj_nyc)

station_ppp_nyc <- ppp(
  x = station_coords_nyc[,1],
  y = station_coords_nyc[,2],
  window = boundary_owin_nyc
)

obs_dist_nyc <- nndist(dock_ppp_nyc, station_ppp_nyc)
obs_mean_dist_nyc <- mean(obs_dist_nyc)


nsim <- 999
sim_mean_dist_nyc <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_nyc <- rpoint(
    n = dock_ppp_nyc$n,
    f = pop_im_nyc,
    win = boundary_owin_nyc
  )
  sim_dist <- nndist(sim_ppp_nyc, station_ppp_nyc)
  sim_mean_dist_nyc[i] <- mean(sim_dist)
}


p_value_nyc <- (sum(sim_mean_dist_nyc <= obs_mean_dist_nyc) + 1) / (nsim + 1)

x_min <- min(sim_mean_dist_nyc, obs_mean_dist_nyc)
x_max <- max(sim_mean_dist_nyc, obs_mean_dist_nyc)

hist(
  sim_mean_dist_nyc,
  main = "New York, NY: Monte Carlo Distribution of\nDistance from Docks to Nearest Train Station",
  xlab = "Simulated Mean Distance",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(x_min, x_max)
)

abline(v = obs_mean_dist_nyc, col = "red", lwd = 3)
p_value_nyc
```

### Boston, MA

```{r}
stations_proj_bos <- st_transform(stations_bos, 32618)
station_coords_bos <- st_coordinates(stations_proj_bos)

station_ppp_bos <- ppp(
  x = station_coords_bos[,1],
  y = station_coords_bos[,2],
  window = boundary_owin_bos
)

obs_dist_bos <- nndist(dock_ppp_bos, station_ppp_bos)
obs_mean_dist_bos <- mean(obs_dist_bos)


nsim <- 999
sim_mean_dist_bos <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_bos <- rpoint(
    n = dock_ppp_bos$n,
    f = pop_im_bos,
    win = boundary_owin_bos
  )
  sim_dist <- nndist(sim_ppp_bos, station_ppp_bos)
  sim_mean_dist_bos[i] <- mean(sim_dist)
}


p_value_bos <- (sum(sim_mean_dist_bos <= obs_mean_dist_bos) + 1) / (nsim + 1)

x_min <- min(sim_mean_dist_bos, obs_mean_dist_bos)
x_max <- max(sim_mean_dist_bos, obs_mean_dist_bos)

hist(
  sim_mean_dist_bos,
  main = "Boston, MA: Monte Carlo Distribution of\nDistance from Docks to Nearest Train Station",
  xlab = "Simulated Mean Distance",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(x_min, x_max)
)

abline(v = obs_mean_dist_bos, col = "red", lwd = 3)
p_value_bos
```

### San Francisco, CA

```{r}
stations_proj_sf <- st_transform(stations_sf, 32618)
station_coords_sf <- st_coordinates(stations_proj_sf)

station_ppp_sf <- ppp(
  x = station_coords_sf[,1],
  y = station_coords_sf[,2],
  window = boundary_owin_sf
)

obs_dist_sf <- nndist(dock_ppp_sf, station_ppp_sf)
obs_mean_dist_sf <- mean(obs_dist_sf)


nsim <- 999
sim_mean_dist_sf <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_sf <- rpoint(
    n = dock_ppp_sf$n,
    f = pop_im_sf,
    win = boundary_owin_sf
  )
  sim_dist <- nndist(sim_ppp_sf, station_ppp_sf)
  sim_mean_dist_sf[i] <- mean(sim_dist)
}


p_value_sf <- (sum(sim_mean_dist_sf <= obs_mean_dist_sf) + 1) / (nsim + 1)

x_min <- min(sim_mean_dist_sf, obs_mean_dist_sf)
x_max <- max(sim_mean_dist_sf, obs_mean_dist_sf)

hist(
  sim_mean_dist_sf,
  main = "San Francisco, CA: Monte Carlo Distribution of\nDistance from Docks to Nearest Train Station",
  xlab = "Simulated Mean Distance",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(x_min, x_max)
)

abline(v = obs_mean_dist_sf, col = "red", lwd = 3)
p_value_sf
```

### Chicago, IL

```{r}
stations_proj_chi <- st_transform(stations_chi, 32618)
station_coords_chi <- st_coordinates(stations_proj_chi)

station_ppp_chi <- ppp(
  x = station_coords_chi[,1],
  y = station_coords_chi[,2],
  window = boundary_owin_chi
)

obs_dist_chi <- nndist(dock_ppp_chi, station_ppp_chi)
obs_mean_dist_chi <- mean(obs_dist_chi)


nsim <- 999
sim_mean_dist_chi <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_chi <- rpoint(
    n = dock_ppp_chi$n,
    f = pop_im_chi,
    win = boundary_owin_chi
  )
  sim_dist <- nndist(sim_ppp_chi, station_ppp_chi)
  sim_mean_dist_chi[i] <- mean(sim_dist)
}


p_value_chi <- (sum(sim_mean_dist_chi <= obs_mean_dist_chi) + 1) / (nsim + 1)

x_min <- min(sim_mean_dist_chi, obs_mean_dist_chi)
x_max <- max(sim_mean_dist_chi, obs_mean_dist_chi)

hist(
  sim_mean_dist_chi,
  main = "Chicago, IL: Monte Carlo Distribution of\nDistance from Docks to Nearest Train Station",
  xlab = "Simulated Mean Distance",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(x_min, x_max)
)

abline(v = obs_mean_dist_chi, col = "red", lwd = 3)
p_value_chi
```
:::

### Monte Carlo Simulation Results against Train Station Locations

| Metropolitan Area | p-value |
|-------------------|---------|
| Washington, DC    | $0.001$ |
| New York, NY      | $1$     |
| Boston, MA        | $0.817$ |
| San Francisco, CA | $0.642$ |
| Chicago, IL       | $1$     |

These results provide the most evidence for diverging strategies for deploying bikeshare docks. We've eliminated population density as a factor by simulating placements that are influenced by population density, a phenomenon confirmed by the first set of hypothesis tests.

In Washington, we can see that simulations give us bikeshare docks that are further away from nearby train stations than the real-life case. Interestingly, this is entirely flipped for New York and Chicago. This tells us that the deployment strategy in Washington deliberately favors those with access to existing public transit, while New York and Chicago provide bikeshare access to areas with less train service.

Boston and San Francisco produced less decisive results, indicating that the presence of public transit may have been less of a consideration for designing those systems. We cannot definitively say that neighborhoods with abundant or lacking public transit are given greater bikeshare access than expected.

## Discussion

Based on the results obtained above, we can conclude that there are multiply strategies for deploying bikeshare docks that result in successful systems with high usage. All five of these cities have far outpaced the rest of the country in producing heavily adopted bikeshare options. What this analysis provides is some insight into how organizers deal with conflicting priorities.

On one hand, bikeshare docks provide a useful supplement to people that prefer car-free travel, who may be using it as an intermediary between longer-distance public transit. On the other hand, providing affordable transportation to those without other car-free options is certainly an appealing prospect. In Washington, we see a system that prioritizes the former, with bikeshare docks aligning strongly with train stations. In New York and Chicago, we see the opposite approach, with bikeshare options following population density to provide widespread access regardless of existing public transit coverage.

Future analysis could be done by finding ways to better isolate specific variables, as well as identifying other potential influences on bikeshare dock placements. Additionally, including cities that have been less successful in implementing bikeshare systems with high usage could provide additional insight into what makes these systems work. Finally, obtaining data on ridership for each dock would be an interesting layer of analysis. To what extent do docks closer to population centers and train stations receive higher usage than the rest?

## References