---
title: Strategic Analysis of Micromobility Dock Deployment
authors:
  - name: Joshua Sweren
    affiliation: Georgetown University
    roles: writing
    corresponding: true
bibliography: references.bib
format:
  html:
    embed-resources: true
    code-fold: true
    echo: true
execute:
  warning: false
  message: false
toc: true
---

```{r}
#install.packages(c("osmdata", "sf", "dplyr", "ggplot2", "rnaturalearth"))
library(osmdata)
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(osmextract)
library(tigris)
library(mapview)
library(lehdr)
library(viridis)
library(viridisLite)
library(spatstat)
library(units)
library(spdep)
library(terra)
library(raster)
library(tictoc)
library(tidycensus)
library(leaflet)
```

## Introduction

Over the last decade, bikeshare systems have become increasingly prevalent in cities around the world in recent years, offering residents and visitors a convenient, low-cost, and environmentally friendly transportation option. However, the spatial deployment of these systems has significant implications over which communities benefit the most and the extent to which they can be of service. This project will analyze the geographic distribution of micromobility stations in several metropolitan areas and evaluate how access to these low-cost transportation options varies across neighborhoods.

As of October 2025, there are five metropolitan areas in the U.S. whose docked bikeshare ridership far exceeds the rest: New York City, Boston, San Francisco, Chicago, and Washington, DC. These cities’ bikeshare systems each boast consistent monthly ridership of over 400,000, while the sixth-place city registers under 100,000. Thus, this project will focus on five cities with proven commitment to developing a system with geographic breadth. Unsurprisingly, these cities also have existing robust public transit networks, so this project will not only look at spatial deployment in relation to population, but also in relation to existing amenities for certain residents. Using OpenStreetMap data to locate bikeshare docks and other transit hubs, along with U.S. Census data to map population density, we will compare how bikeshare has been spatially deployed in its most prominent cases.

## Literature Review

## Methodology

Each city will be analyzed via the following sets of hypotheses:

1.  Docks in relation to population density:

    -   **Null** **hypothesis**: There is no relationship between the spatial distribution of bikeshare docks and population density. Dock locations occur independently of areas with higher or lower population density.

    <!-- -->

    -   **Alternative** **hypothesis**: There is a relationship between the spatial distribution of Capital Bikeshare docks and population density. Dock locations are more likely to occur in areas with higher population density.

2.  Docks in relation to other public transit options:

    -   **Null hypothesis**: There is no relationship between the spatial distribution of bikeshare docks and the presence of other public transit nodes.

    -   **Alternative hypothesis:** There is a relationship between the spatial distribution of bikeshare docks and the presence of other public transit nodes. Dock locations are more likely to be placed in areas with existing public transit infrastructure.

The first steps towards testing these hypotheses will be to gather bikeshare dock and train station locations from OpenStreetMap via the `osm` package in R and population density data from the U.S. Census via the `tidycensus` package. To create a geographic region for each city, we will create a concave hull around all of the bikeshare docks. Since metropolitan areas differ in the distance from the city's center that bikeshare systems reach, this will allow us to only consider areas within the chosen scope of each system, and how docks have been deployed within those limitations.

The following information will then be gathered to test the hypotheses:

-   Moran's I to generally understand clustering of bike docks.

-   Regression testing results to analyze the relationship between dock locations, population density, and public transit options.

-   Monte Carlo simulations to compare actual results to randomized data.

## Exploratory Data Analysis

::: panel-tabset
### Washington, DC

```{r}
dc_bb <- c(xmin = -77.45, ymin = 38.70, xmax = -76.80, ymax = 39.10)
docks_raw_dc <- opq(bbox = dc_bb) |>
  add_osm_feature(key = "amenity", value = "bicycle_rental") |>
  add_osm_feature(key = "network", value = "Capital Bikeshare") |>
  osmdata_sf()

docks_dc <- docks_raw_dc$osm_points |>
  st_as_sf() |>
  st_transform(3857) %>%
  filter(!is.na(network))
```

```{r}
hull_dc <- docks_dc |> 
    st_union() |> 
    st_concave_hull(.2)
hull_buffered_dc <- st_buffer(hull_dc, dist = 2500)
hull_buffered_wgs84_dc <- st_transform(hull_buffered_dc, 4326)
docks_dc <- st_intersection(
    st_transform(docks_dc, 4326), 
    hull_buffered_wgs84_dc)
```

```{r}
mapview(hull_buffered_wgs84_dc, col.regions = "lightblue", alpha.regions = 0.3) +
  mapview(docks_dc, col.regions = "red")
```

### New York, NY

```{r}
nyc_bb <- c(xmin = -74.15, ymin = 40.55, xmax = -73.70, ymax = 40.95)
docks_raw_nyc <- opq(bbox = nyc_bb) |>
  add_osm_feature(key = "amenity", value = "bicycle_rental") |>
  add_osm_feature(key = "network", value = "Citi Bike") |>
  osmdata_sf()

docks_nyc <- docks_raw_nyc$osm_points |>
  st_as_sf() |>
  st_transform(3857) %>%
  filter(!is.na(network))
```

```{r}
hull_nyc <- docks_nyc |> 
  st_union() |> 
  st_concave_hull(.2)
hull_buffered_nyc <- st_buffer(hull_nyc, dist = 500)
hull_buffered_wgs84_nyc <- st_transform(hull_buffered_nyc, 4326)
docks_nyc <- st_intersection(
  st_transform(docks_nyc, 4326),
  hull_buffered_wgs84_nyc
)
```

```{r}
mapview(hull_buffered_wgs84_nyc, col.regions = "lightblue", alpha.regions = 0.3) +
  mapview(docks_nyc, col.regions = "red")
```

### Boston, MA

```{r}
bos_bb <- c(-71.20, 42.20, -70.90, 42.45)
docks_raw_bos <- opq(bbox = bos_bb) |>
  add_osm_feature(key = "amenity", value = "bicycle_rental") |>
  add_osm_feature(key = "network", value = "Bluebikes") |>
  osmdata_sf()

docks_bos <- docks_raw_bos$osm_points |>
  st_as_sf() |>
  st_transform(3857) %>%
  filter(!is.na(network))
```

```{r}
hull_bos <- docks_bos |> 
  st_union() |> 
  st_concave_hull(.2)
hull_buffered_bos <- st_buffer(hull_bos, dist = 1000)
hull_buffered_wgs84_bos <- st_transform(hull_buffered_bos, 4326)
docks_bos <- st_intersection(
  st_transform(docks_bos, 4326),
  hull_buffered_wgs84_bos
)
```

```{r}
mapview(hull_buffered_wgs84_bos, col.regions = "lightblue", alpha.regions = 0.3) +
  mapview(docks_bos, col.regions = "red")
```

### San Francisco, CA

```{r}
sf_bb <- c(-122.55, 37.70, -122.35, 37.85)
docks_raw_sf <- opq(bbox = sf_bb) |>
  add_osm_feature(key = "amenity", value = "bicycle_rental") |>
  add_osm_feature(key = "network", value = "Bay Wheels") |>
  osmdata_sf()

docks_sf <- docks_raw_sf$osm_points |>
  st_as_sf() |>
  st_transform(3857) %>%
  filter(!is.na(network))
```

```{r}
hull_sf <- docks_sf |> 
  st_union() |> 
  st_concave_hull(.2)
hull_buffered_sf <- st_buffer(hull_sf, dist = 1000)
hull_buffered_wgs84_sf <- st_transform(hull_buffered_sf, 4326)
docks_sf <- st_intersection(
  st_transform(docks_sf, 4326),
  hull_buffered_wgs84_sf
)
```

```{r}
mapview(hull_buffered_wgs84_sf, col.regions = "lightblue", alpha.regions = 0.3) +
  mapview(docks_sf, col.regions = "red")
```

### Chicago, IL

```{r}
chi_bb <- c(-87.80, 41.75, -87.55, 42.00)
docks_raw_chi <- opq(bbox = chi_bb) |>
  add_osm_feature(key = "amenity", value = "bicycle_rental") |>
  add_osm_feature(key = "network", value = "Divvy") |>
  osmdata_sf()

docks_chi <- docks_raw_chi$osm_points |>
  st_as_sf() |>
  st_transform(3857) %>%
  filter(!is.na(network))
```

```{r}
hull_chi <- docks_chi |> 
  st_union() |> 
  st_concave_hull(.2)
hull_buffered_chi <- st_buffer(hull_chi, dist = 2500)
hull_buffered_wgs84_chi <- st_transform(hull_buffered_chi, 4326)
docks_chi <- st_intersection(
  st_transform(docks_chi, 4326),
  hull_buffered_wgs84_chi
)
```

```{r}
mapview(hull_buffered_wgs84_chi, col.regions = "lightblue", alpha.regions = 0.3) +
  mapview(docks_chi, col.regions = "red")
```
:::

## Plotting Other Data

::: panel-tabset
### Washington, DC

```{r}
#census_api_key("5e8593d6e289e6ae04c33a019f4951ca2bdd9523", install = TRUE)
dc_bg <- get_acs(
  geography = "block group",
  variables = "B01003_001E",
  year = 2022,
  survey = "acs5",
  state = c("DC", "MD", "VA"),
  geometry = TRUE
)
```

```{r}
dc_bg <- st_transform(dc_bg, st_crs(hull_buffered_wgs84_dc))

dc_bg <- dc_bg %>%
  mutate(area_original = as.numeric(st_area(geometry)))

dc_bg_clipped <- st_intersection(dc_bg, hull_buffered_wgs84_dc)

dc_bg_clipped <- dc_bg_clipped %>%
  mutate(
    area_clipped = as.numeric(st_area(geometry)),
    pop_adjusted = estimate * (area_clipped / area_original),
    area_km2 = area_clipped / 1e6,
    pop_density = pop_adjusted / area_km2
  )

dc_bg_density <- dc_bg_clipped %>%
  mutate(
    area_km2 = as.numeric(st_area(geometry)) / 1e6,
    pop_density = estimate / area_km2
  )
```

```{r}
max_density <- 20000
dc_bg_density$pop_density_plot <- pmin(dc_bg_density$pop_density, max_density)
```

```{r}
dc_bbox <- st_bbox(hull_buffered_wgs84_dc)

# bus_stops <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "bus_stop") %>%
#   osmdata_sf() %>%
#   .$osm_points %>%
#   st_transform(st_crs(hull_buffered_wgs84_dc))

stations_dc <- opq(dc_bbox) %>%
  add_osm_feature(key = "railway", value = "station") %>%
  osmdata_sf() %>%
  .$osm_points %>%
  st_transform(st_crs(hull_buffered_wgs84_dc))
```

```{r}
pal <- leaflet::colorNumeric(viridis(20), domain = dc_bg_density$pop_density_plot)
leaflet() %>%
  leaflet::addProviderTiles("CartoDB.Positron") %>%
  
  # Population density polygons
  leaflet::addPolygons(
    data = dc_bg_density,
    fillColor = ~pal(pop_density_plot),
    fillOpacity = 0.6,
    color = "grey",
    weight = 0.5,
    group = "Population Density",
    popup = ~paste0("Population density: ", round(pop_density_plot))
  ) %>%
  
  # Dock locations
  leaflet::addCircleMarkers(
    data = docks_dc,
    radius = 1,
    color = "blue",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Bikeshare Docks"
  ) %>%
  
  # Train stations
  leaflet::addCircleMarkers(
    data = stations_dc,
    radius = 1,
    color = "red",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Train Stations"
  ) %>%
  
  # Layer controls
  leaflet::addLayersControl(
    overlayGroups = c("Population Density", "Bikeshare Docks", "Train Stations"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
```

### New York, NY

```{r}
nyc_bg <- get_acs(
  geography = "block group",
  variables = "B01003_001E",
  year = 2022,
  survey = "acs5",
  state = c("NY", "NJ"),
  geometry = TRUE
)
```

```{r}
nyc_bg <- st_transform(nyc_bg, st_crs(hull_buffered_wgs84_nyc))

nyc_bg <- nyc_bg %>%
  mutate(area_original = as.numeric(st_area(geometry)))

nyc_bg_clipped <- st_intersection(nyc_bg, hull_buffered_wgs84_nyc)

nyc_bg_clipped <- nyc_bg_clipped %>%
  mutate(
    area_clipped = as.numeric(st_area(geometry)),
    pop_adjusted = estimate * (area_clipped / area_original),
    area_km2 = area_clipped / 1e6,
    pop_density = pop_adjusted / area_km2
  )

nyc_bg_density <- nyc_bg_clipped %>%
  mutate(
    area_km2 = as.numeric(st_area(geometry)) / 1e6,
    pop_density = estimate / area_km2
  )
```

```{r}
max_density <- 75000
nyc_bg_density$pop_density_plot <- pmin(nyc_bg_density$pop_density, max_density)
```

```{r}
nyc_bbox <- st_bbox(hull_buffered_wgs84_nyc)

# bus_stops <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "bus_stop") %>%
#   osmdata_sf() %>%
#   .$osm_points %>%
#   st_transform(st_crs(hull_buffered_wgs84_dc))

stations_nyc <- opq(nyc_bbox) %>%
  add_osm_feature(key = "railway", value = "station") %>%
  osmdata_sf() %>%
  .$osm_points %>%
  st_transform(st_crs(hull_buffered_wgs84_nyc))
```

```{r}
pal <- leaflet::colorNumeric(viridis(20), domain = nyc_bg_density$pop_density_plot)
leaflet() %>%
  leaflet::addProviderTiles("CartoDB.Positron") %>%
  
  # Population density polygons
  leaflet::addPolygons(
    data = nyc_bg_density,
    fillColor = ~pal(pop_density_plot),
    fillOpacity = 0.6,
    color = "grey",
    weight = 0.5,
    group = "Population Density",
    popup = ~paste0("Population density: ", round(pop_density_plot))
  ) %>%
  
  # Dock locations
  leaflet::addCircleMarkers(
    data = docks_nyc,
    radius = 1,
    color = "blue",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Bikeshare Docks"
  ) %>%
  
  # Train stations
  leaflet::addCircleMarkers(
    data = stations_nyc,
    radius = 1,
    color = "red",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Train Stations"
  ) %>%
  
  # Layer controls
  leaflet::addLayersControl(
    overlayGroups = c("Population Density", "Bikeshare Docks", "Train Stations"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
```

### Boston, MA

```{r}
bos_bg <- get_acs(
  geography = "block group",
  variables = "B01003_001E",
  year = 2022,
  survey = "acs5",
  state = c("MA"),
  geometry = TRUE
)
```

```{r}
bos_bg <- st_transform(bos_bg, st_crs(hull_buffered_wgs84_bos))

bos_bg <- bos_bg %>%
  mutate(area_original = as.numeric(st_area(geometry)))

bos_bg_clipped <- st_intersection(bos_bg, hull_buffered_wgs84_bos)

bos_bg_clipped <- bos_bg_clipped %>%
  mutate(
    area_clipped = as.numeric(st_area(geometry)),
    pop_adjusted = estimate * (area_clipped / area_original),
    area_km2 = area_clipped / 1e6,
    pop_density = pop_adjusted / area_km2
  )

bos_bg_density <- bos_bg_clipped %>%
  mutate(
    area_km2 = as.numeric(st_area(geometry)) / 1e6,
    pop_density = estimate / area_km2
  )
```

```{r}
max_density <- 25000
bos_bg_density$pop_density_plot <- pmin(bos_bg_density$pop_density, max_density)
```

```{r}
bos_bbox <- st_bbox(hull_buffered_wgs84_bos)

# bus_stops <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "bus_stop") %>%
#   osmdata_sf() %>%
#   .$osm_points %>%
#   st_transform(st_crs(hull_buffered_wgs84_dc))

stations_bos <- opq(bos_bbox) %>%
  add_osm_feature(key = "railway", value = "station") %>%
  osmdata_sf() %>%
  .$osm_points %>%
  st_transform(st_crs(hull_buffered_wgs84_bos))
```

```{r}
pal <- leaflet::colorNumeric(viridis(20), domain = bos_bg_density$pop_density_plot)
leaflet() %>%
  leaflet::addProviderTiles("CartoDB.Positron") %>%
  
  # Population density polygons
  leaflet::addPolygons(
    data = bos_bg_density,
    fillColor = ~pal(pop_density_plot),
    fillOpacity = 0.6,
    color = "grey",
    weight = 0.5,
    group = "Population Density",
    popup = ~paste0("Population density: ", round(pop_density_plot))
  ) %>%
  
  # Dock locations
  leaflet::addCircleMarkers(
    data = docks_bos,
    radius = 1,
    color = "blue",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Bikeshare Docks"
  ) %>%
  
  # Train stations
  leaflet::addCircleMarkers(
    data = stations_bos,
    radius = 1,
    color = "red",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Train Stations"
  ) %>%
  
  # Layer controls
  leaflet::addLayersControl(
    overlayGroups = c("Population Density", "Bikeshare Docks", "Train Stations"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
```

### San Francisco, CA

```{r}
sf_bg <- get_acs(
  geography = "block group",
  variables = "B01003_001E",
  year = 2022,
  survey = "acs5",
  state = c("CA"),
  geometry = TRUE
)
```

```{r}
sf_bg <- st_transform(sf_bg, st_crs(hull_buffered_wgs84_sf))

sf_bg <- sf_bg %>%
  mutate(area_original = as.numeric(st_area(geometry)))

sf_bg_clipped <- st_intersection(sf_bg, hull_buffered_wgs84_sf)

sf_bg_clipped <- sf_bg_clipped %>%
  mutate(
    area_clipped = as.numeric(st_area(geometry)),
    pop_adjusted = estimate * (area_clipped / area_original),
    area_km2 = area_clipped / 1e6,
    pop_density = pop_adjusted / area_km2
  )

sf_bg_density <- sf_bg_clipped %>%
  mutate(
    area_km2 = as.numeric(st_area(geometry)) / 1e6,
    pop_density = estimate / area_km2
  )
```

```{r}
max_density <- 35000
sf_bg_density$pop_density_plot <- pmin(sf_bg_density$pop_density, max_density)
```

```{r}
sf_bbox <- st_bbox(hull_buffered_wgs84_sf)

# bus_stops <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "bus_stop") %>%
#   osmdata_sf() %>%
#   .$osm_points %>%
#   st_transform(st_crs(hull_buffered_wgs84_dc))

stations_sf <- opq(sf_bbox) %>%
  add_osm_feature(key = "railway", value = "station") %>%
  osmdata_sf() %>%
  .$osm_points %>%
  st_transform(st_crs(hull_buffered_wgs84_sf))
```

```{r}
pal <- leaflet::colorNumeric(viridis(20), domain = sf_bg_density$pop_density_plot)
leaflet() %>%
  leaflet::addProviderTiles("CartoDB.Positron") %>%
  
  # Population density polygons
  leaflet::addPolygons(
    data = sf_bg_density,
    fillColor = ~pal(pop_density_plot),
    fillOpacity = 0.6,
    color = "grey",
    weight = 0.5,
    group = "Population Density",
    popup = ~paste0("Population density: ", round(pop_density_plot))
  ) %>%
  
  # Dock locations
  leaflet::addCircleMarkers(
    data = docks_sf,
    radius = 1,
    color = "blue",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Bikeshare Docks"
  ) %>%
  
  # Train stations
  leaflet::addCircleMarkers(
    data = stations_sf,
    radius = 1,
    color = "red",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Train Stations"
  ) %>%
  
  # Layer controls
  leaflet::addLayersControl(
    overlayGroups = c("Population Density", "Bikeshare Docks", "Train Stations"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
```

### Chicago, IL

```{r}
chi_bg <- get_acs(
  geography = "block group",
  variables = "B01003_001E",
  year = 2022,
  survey = "acs5",
  state = c("IL"),
  geometry = TRUE
)
```

```{r}
chi_bg <- st_transform(chi_bg, st_crs(hull_buffered_wgs84_chi))

chi_bg <- chi_bg %>%
  mutate(area_original = as.numeric(st_area(geometry)))

chi_bg_clipped <- st_intersection(chi_bg, hull_buffered_wgs84_chi)

chi_bg_clipped <- chi_bg_clipped %>%
  mutate(
    area_clipped = as.numeric(st_area(geometry)),
    pop_adjusted = estimate * (area_clipped / area_original),
    area_km2 = area_clipped / 1e6,
    pop_density = pop_adjusted / area_km2
  )

chi_bg_density <- chi_bg_clipped %>%
  mutate(
    area_km2 = as.numeric(st_area(geometry)) / 1e6,
    pop_density = estimate / area_km2
  )
```

```{r}
max_density <- 30000
chi_bg_density$pop_density_plot <- pmin(chi_bg_density$pop_density, max_density)
```

```{r}
chi_bbox <- st_bbox(hull_buffered_wgs84_chi)

# bus_stops <- opq(bbox) %>%
#   add_osm_feature(key = "highway", value = "bus_stop") %>%
#   osmdata_sf() %>%
#   .$osm_points %>%
#   st_transform(st_crs(hull_buffered_wgs84_dc))

stations_chi <- opq(chi_bbox) %>%
  add_osm_feature(key = "railway", value = "station") %>%
  osmdata_sf() %>%
  .$osm_points %>%
  st_transform(st_crs(hull_buffered_wgs84_chi))
```

```{r}
pal <- leaflet::colorNumeric(viridis(20), domain = chi_bg_density$pop_density_plot)
leaflet() %>%
  leaflet::addProviderTiles("CartoDB.Positron") %>%
  
  # Population density polygons
  leaflet::addPolygons(
    data = chi_bg_density,
    fillColor = ~pal(pop_density_plot),
    fillOpacity = 0.6,
    color = "grey",
    weight = 0.5,
    group = "Population Density",
    popup = ~paste0("Population density: ", round(pop_density_plot))
  ) %>%
  
  # Dock locations
  leaflet::addCircleMarkers(
    data = docks_chi,
    radius = 1,
    color = "blue",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Bikeshare Docks"
  ) %>%
  
  # Train stations
  leaflet::addCircleMarkers(
    data = stations_chi,
    radius = 1,
    color = "red",
    fill = TRUE,
    fillOpacity = 0.5,
    group = "Train Stations"
  ) %>%
  
  # Layer controls
  leaflet::addLayersControl(
    overlayGroups = c("Population Density", "Bikeshare Docks", "Train Stations"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
```
:::

## Univariate Testing

::: panel-tabset
### Washington, DC

```{r}
coords <- st_coordinates(docks_dc)

knn <- knearneigh(coords, k = 4)
nb <- knn2nb(knn)

lw <- nb2listw(nb, style = "W")

inv_dist <- sapply(1:nrow(coords), function(i) {
  neighbors <- nb[[i]]
  if(length(neighbors) == 0) return(0)
  1 / mean(sqrt((coords[i,1]-coords[neighbors,1])^2 + (coords[i,2]-coords[neighbors,2])^2))
})

moran_result <- moran.test(inv_dist, lw, zero.policy = TRUE)
moran_result
```

```{r}
dock_values <- inv_dist

localI <- localmoran(dock_values, lw, zero.policy = TRUE)
docks_dc$localI <- localI[,1]

leaflet(docks_dc) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    radius = 4,
    color = ~viridis::viridis(20)[cut(localI, breaks = 20)],
    fillOpacity = 0.8,
    popup = ~paste0("Local Moran's I: ", round(localI, 3))
  )
```

### New York, NY

```{r}
coords <- st_coordinates(docks_nyc)

knn <- knearneigh(coords, k = 4)
nb <- knn2nb(knn)

lw <- nb2listw(nb, style = "W")

inv_dist <- sapply(1:nrow(coords), function(i) {
  neighbors <- nb[[i]]
  if(length(neighbors) == 0) return(0)
  1 / mean(sqrt((coords[i,1]-coords[neighbors,1])^2 + (coords[i,2]-coords[neighbors,2])^2))
})

moran_result <- moran.test(inv_dist, lw, zero.policy = TRUE)
moran_result
```

```{r}
dock_values <- inv_dist

localI <- localmoran(dock_values, lw, zero.policy = TRUE)
docks_nyc$localI <- localI[,1]

leaflet(docks_nyc) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    radius = 4,
    color = ~viridis::viridis(20)[cut(localI, breaks = 20)],
    fillOpacity = 0.8,
    popup = ~paste0("Local Moran's I: ", round(localI, 3))
  )
```

### Boston, MA

```{r}
coords <- st_coordinates(docks_bos)

knn <- knearneigh(coords, k = 4)
nb <- knn2nb(knn)

lw <- nb2listw(nb, style = "W")

inv_dist <- sapply(1:nrow(coords), function(i) {
  neighbors <- nb[[i]]
  if(length(neighbors) == 0) return(0)
  1 / mean(sqrt((coords[i,1]-coords[neighbors,1])^2 + (coords[i,2]-coords[neighbors,2])^2))
})

moran_result <- moran.test(inv_dist, lw, zero.policy = TRUE)
moran_result
```

```{r}
dock_values <- inv_dist

localI <- localmoran(dock_values, lw, zero.policy = TRUE)
docks_bos$localI <- localI[,1]

leaflet(docks_bos) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    radius = 4,
    color = ~viridis::viridis(20)[cut(localI, breaks = 20)],
    fillOpacity = 0.8,
    popup = ~paste0("Local Moran's I: ", round(localI, 3))
  )
```

### San Francisco, CA

```{r}
coords <- st_coordinates(docks_sf)

knn <- knearneigh(coords, k = 4)
nb <- knn2nb(knn)

lw <- nb2listw(nb, style = "W")

inv_dist <- sapply(1:nrow(coords), function(i) {
  neighbors <- nb[[i]]
  if(length(neighbors) == 0) return(0)
  1 / mean(sqrt((coords[i,1]-coords[neighbors,1])^2 + (coords[i,2]-coords[neighbors,2])^2))
})

moran_result <- moran.test(inv_dist, lw, zero.policy = TRUE)
moran_result
```

```{r}
dock_values <- inv_dist

localI <- localmoran(dock_values, lw, zero.policy = TRUE)
docks_sf$localI <- localI[,1]

leaflet(docks_sf) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    radius = 4,
    color = ~viridis::viridis(20)[cut(localI, breaks = 20)],
    fillOpacity = 0.8,
    popup = ~paste0("Local Moran's I: ", round(localI, 3))
  )
```

### Chicago, IL

```{r}
coords <- st_coordinates(docks_chi)

knn <- knearneigh(coords, k = 4)
nb <- knn2nb(knn)

lw <- nb2listw(nb, style = "W")

inv_dist <- sapply(1:nrow(coords), function(i) {
  neighbors <- nb[[i]]
  if(length(neighbors) == 0) return(0)
  1 / mean(sqrt((coords[i,1]-coords[neighbors,1])^2 + (coords[i,2]-coords[neighbors,2])^2))
})

moran_result <- moran.test(inv_dist, lw, zero.policy = TRUE)
moran_result
```

```{r}
dock_values <- inv_dist

localI <- localmoran(dock_values, lw, zero.policy = TRUE)
docks_chi$localI <- localI[,1]

leaflet(docks_chi) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(
    radius = 4,
    color = ~viridis::viridis(20)[cut(localI, breaks = 20)],
    fillOpacity = 0.8,
    popup = ~paste0("Local Moran's I: ", round(localI, 3))
  )
```
:::

### Univariate Results

| Metropolitan Area | Moran's I Statistic |
|-------------------|---------------------|
| Washington, DC    | $0.8444759043$      |
| New York, NY      | $0.6827748494$      |
| Boston, MA        | $0.842640558$       |
| San Francisco, CA | $0.808442476$       |
| Chicago, IL       | $0.9584339511$      |

## Hypothesis Test: Population Density

### Correlation Functions

::: panel-tabset
### Washington, DC

```{r}
proj_crs <- "EPSG:32618"
docks_proj_dc <- st_transform(docks_dc, 32618)
boundary_proj_dc <- st_transform(hull_buffered_wgs84_dc, 32618)
boundary_owin_dc <- as.owin(boundary_proj_dc)

dock_coords_dc <- st_coordinates(docks_proj_dc)

dock_ppp_dc <- ppp(
  x = dock_coords_dc[,1],
  y = dock_coords_dc[,2],
  window = boundary_owin_dc
)
dc_bg_density_proj <- st_transform(dc_bg_density, proj_crs)
rtemplate_dc <- rast(
  extent = ext(dc_bg_density_proj),
  resolution = 250,
  crs = "EPSG:32618"
)

pop_rast_dc <- rasterize(
  vect(dc_bg_density_proj),
  rtemplate_dc,
  field = "pop_density_plot",
  fun = "mean"
)

pop_rast_dc <- mask(pop_rast_dc, vect(boundary_proj_dc))
```

```{r}
nx_dc <- ncol(pop_rast_dc)
ny_dc <- nrow(pop_rast_dc)
r_ext_dc <- ext(pop_rast_dc)
res_x_dc <- res(pop_rast_dc)[1]
res_y_dc <- res(pop_rast_dc)[2]

xcol_dc <- seq(r_ext_dc$xmin + res_x_dc/2, r_ext_dc$xmax - res_x_dc/2, length.out = nx_dc)
yrow_dc <- seq(r_ext_dc$ymin + res_y_dc/2, r_ext_dc$ymax - res_y_dc/2, length.out = ny_dc)

vals_dc <- values(pop_rast_dc, mat = FALSE)
mat_dc <- matrix(vals_dc, nrow = ny_dc, ncol = nx_dc, byrow = TRUE)

mat_oriented_dc <- mat_dc[ny_dc:1, , drop = FALSE]
pop_im_dc <- im(mat_oriented_dc, xcol_dc, yrow_dc)
```

```{r}
m0_dc <- ppm(dock_ppp_dc ~ 1)
m1_dc <- ppm(dock_ppp_dc ~ pop_im_dc)
summary(m1_dc)
anova(m0_dc, m1_dc, test="LR")
```

### New York, NY

```{r}
docks_proj_nyc <- st_transform(docks_nyc, 32618)
boundary_proj_nyc <- st_transform(hull_buffered_wgs84_nyc, 32618)
boundary_owin_nyc <- as.owin(boundary_proj_nyc)

dock_coords_nyc <- st_coordinates(docks_proj_nyc)

dock_ppp_nyc <- ppp(
  x = dock_coords_nyc[,1],
  y = dock_coords_nyc[,2],
  window = boundary_owin_nyc
)
nyc_bg_density_proj <- st_transform(nyc_bg_density, crs = proj_crs)

rtemplate_nyc <- rast(
  extent = ext(nyc_bg_density_proj),
  resolution = 250,
  crs = "EPSG:32618"
)

pop_rast_nyc <- rasterize(
  vect(nyc_bg_density_proj),
  rtemplate_nyc,
  field = "pop_density_plot",
  fun = "mean"
)
pop_rast_nyc <- mask(pop_rast_nyc, vect(boundary_proj_nyc))
```

```{r}
nx_nyc <- ncol(pop_rast_nyc)
ny_nyc <- nrow(pop_rast_nyc)
r_ext_nyc <- ext(pop_rast_nyc)
res_x_nyc <- res(pop_rast_nyc)[1]
res_y_nyc <- res(pop_rast_nyc)[2]

xcol_nyc <- seq(r_ext_nyc$xmin + res_x_nyc/2, r_ext_nyc$xmax - res_x_nyc/2, length.out = nx_nyc)
yrow_nyc <- seq(r_ext_nyc$ymin + res_y_nyc/2, r_ext_nyc$ymax - res_y_nyc/2, length.out = ny_nyc)

vals_nyc <- values(pop_rast_nyc, mat = FALSE)
mat_nyc <- matrix(vals_nyc, nrow = ny_nyc, ncol = nx_nyc, byrow = TRUE)

mat_oriented_nyc <- mat_nyc[ny_nyc:1, , drop = FALSE]
pop_im_nyc <- im(mat_oriented_nyc, xcol_nyc, yrow_nyc)
```

```{r}
m0_nyc <- ppm(dock_ppp_nyc ~ 1)
m1_nyc <- ppm(dock_ppp_nyc ~ pop_im_nyc)
summary(m1_nyc)
anova(m0_nyc, m1_nyc, test="LR")
```

### Boston, MA

```{r}
docks_proj_bos <- st_transform(docks_bos, 32618)
boundary_proj_bos <- st_transform(hull_buffered_wgs84_bos, 32618)
boundary_owin_bos <- as.owin(boundary_proj_bos)

dock_coords_bos <- st_coordinates(docks_proj_bos)

dock_ppp_bos <- ppp(
  x = dock_coords_bos[,1],
  y = dock_coords_bos[,2],
  window = boundary_owin_bos
)
bos_bg_density_proj <- st_transform(bos_bg_density, crs = proj_crs)

rtemplate_bos <- rast(
  extent = ext(bos_bg_density_proj),
  resolution = 250,
  crs = "EPSG:32618"
)

pop_rast_bos <- rasterize(
  vect(bos_bg_density_proj),
  rtemplate_bos,
  field = "pop_density_plot",
  fun = "mean"
)
pop_rast_bos <- mask(pop_rast_bos, vect(boundary_proj_bos))
```

```{r}
nx_bos <- ncol(pop_rast_bos)
ny_bos <- nrow(pop_rast_bos)
r_ext_bos <- ext(pop_rast_bos)
res_x_bos <- res(pop_rast_bos)[1]
res_y_bos <- res(pop_rast_bos)[2]

xcol_bos <- seq(r_ext_bos$xmin + res_x_bos/2, r_ext_bos$xmax - res_x_bos/2, length.out = nx_bos)
yrow_bos <- seq(r_ext_bos$ymin + res_y_bos/2, r_ext_bos$ymax - res_y_bos/2, length.out = ny_bos)

vals_bos <- values(pop_rast_bos, mat = FALSE)
mat_bos <- matrix(vals_bos, nrow = ny_bos, ncol = nx_bos, byrow = TRUE)

mat_oriented_bos <- mat_bos[ny_bos:1, , drop = FALSE]
pop_im_bos <- im(mat_oriented_bos, xcol_bos, yrow_bos)
```

```{r}
m0_bos <- ppm(dock_ppp_bos ~ 1)
m1_bos <- ppm(dock_ppp_bos ~ pop_im_bos)
summary(m1_bos)
anova(m0_bos, m1_bos, test="LR")
```

### San Francisco, CA

```{r}
docks_proj_sf <- st_transform(docks_sf, 32618)
boundary_proj_sf <- st_transform(hull_buffered_wgs84_sf, 32618)
boundary_owin_sf <- as.owin(boundary_proj_sf)

dock_coords_sf <- st_coordinates(docks_proj_sf)

dock_ppp_sf <- ppp(
  x = dock_coords_sf[,1],
  y = dock_coords_sf[,2],
  window = boundary_owin_sf
)
sf_bg_density_proj <- st_transform(sf_bg_density, crs = proj_crs)

rtemplate_sf <- rast(
  extent = ext(sf_bg_density_proj),
  resolution = 250,
  crs = "EPSG:32618"
)

pop_rast_sf <- rasterize(
  vect(sf_bg_density_proj),
  rtemplate_sf,
  field = "pop_density_plot",
  fun = "mean"
)
pop_rast_sf <- mask(pop_rast_sf, vect(boundary_proj_sf))
```

```{r}
nx_sf <- ncol(pop_rast_sf)
ny_sf <- nrow(pop_rast_sf)
r_ext_sf <- ext(pop_rast_sf)
res_x_sf <- res(pop_rast_sf)[1]
res_y_sf <- res(pop_rast_sf)[2]

xcol_sf <- seq(r_ext_sf$xmin + res_x_sf/2, r_ext_sf$xmax - res_x_sf/2, length.out = nx_sf)
yrow_sf <- seq(r_ext_sf$ymin + res_y_sf/2, r_ext_sf$ymax - res_y_sf/2, length.out = ny_sf)

vals_sf <- values(pop_rast_sf, mat = FALSE)
mat_sf <- matrix(vals_sf, nrow = ny_sf, ncol = nx_sf, byrow = TRUE)

mat_oriented_sf <- mat_sf[ny_sf:1, , drop = FALSE]
pop_im_sf <- im(mat_oriented_sf, xcol_sf, yrow_sf)
```

```{r}
m0_sf <- ppm(dock_ppp_sf ~ 1)
m1_sf <- ppm(dock_ppp_sf ~ pop_im_sf)
summary(m1_sf)
anova(m0_sf, m1_sf, test="LR")
```

### Chicago, IL

```{r}
docks_proj_chi <- st_transform(docks_chi, 32618)
boundary_proj_chi <- st_transform(hull_buffered_wgs84_chi, 32618)
boundary_owin_chi <- as.owin(boundary_proj_chi)

dock_coords_chi <- st_coordinates(docks_proj_chi)

dock_ppp_chi <- ppp(
  x = dock_coords_chi[,1],
  y = dock_coords_chi[,2],
  window = boundary_owin_chi
)
chi_bg_density_proj <- st_transform(chi_bg_density, crs = proj_crs)

rtemplate_chi <- rast(
  extent = ext(chi_bg_density_proj),
  resolution = 250,
  crs = "EPSG:32618"
)

pop_rast_chi <- rasterize(
  vect(chi_bg_density_proj),
  rtemplate_chi,
  field = "pop_density_plot",
  fun = "mean"
)
pop_rast_chi <- mask(pop_rast_chi, vect(boundary_proj_chi))
```

```{r}
nx_chi <- ncol(pop_rast_chi)
ny_chi <- nrow(pop_rast_chi)
r_ext_chi <- ext(pop_rast_chi)
res_x_chi <- res(pop_rast_chi)[1]
res_y_chi <- res(pop_rast_chi)[2]

xcol_chi <- seq(r_ext_chi$xmin + res_x_chi/2, r_ext_chi$xmax - res_x_chi/2, length.out = nx_chi)
yrow_chi <- seq(r_ext_chi$ymin + res_y_chi/2, r_ext_chi$ymax - res_y_chi/2, length.out = ny_chi)

vals_chi <- values(pop_rast_chi, mat = FALSE)
mat_chi <- matrix(vals_chi, nrow = ny_chi, ncol = nx_chi, byrow = TRUE)

mat_oriented_chi <- mat_chi[ny_chi:1, , drop = FALSE]
pop_im_chi <- im(mat_oriented_chi, xcol_chi, yrow_chi)
```

```{r}
m0_chi <- ppm(dock_ppp_chi ~ 1)
m1_chi <- ppm(dock_ppp_chi ~ pop_im_chi)
summary(m1_chi)
anova(m0_chi, m1_chi, test="LR")
```
:::

### Correlation Function Results against Population Density

| Metropolitan Area | Docks | Dock Intensity (points/unit squared) | `pop_im` coefficient | Z-value | Deviance diff |
|------------|------------|------------|------------|------------|------------|
| Washington, DC | $774$ | $5.06 \times 10^{-7}$ | $1.913 \times 10^{-4}$ | $33.74$ | $683.18$ |
| New York, NY | $2,303$ | $1.5 \times 10^{-5}$ | $2.121 \times 10^-5$ | $22.44$ | $436.56$ |
| Boston, MA | $320$ | $1.08 \times 10^{-5}$ | $9.189 \times 10^{-5}$ | $11.41$ | $105.4$ |
| San Francisco, CA | $107$ | $6.35 \times 10^{-6}$ | $3.904 \times 10^{-5}$ | $2.60$ | $6.436$ |
| Chicago, IL | $909$ | $3.77 \times 10^{-6}$ | $9.194 \times 10^{-5}$ | $19.56$ | $269.89$ |

### Monte Carlo Simulations

::: panel-tabset
### Washington, DC

```{r}
model_obs_dc <- ppm(dock_ppp_dc ~ pop_im_dc)
coef_obs_dc <- coef(model_obs_dc)["pop_im_dc"]

nsim <- 999
coef_sim_dc <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_dc <- runifpoint(n = dock_ppp_dc$n, win = boundary_owin_dc)
  sim_model_dc <- ppm(sim_ppp_dc ~ pop_im_dc)
  coef_sim_dc[i] <- coef(sim_model_dc)["pop_im_dc"]
}

p_value_dc <- (sum(coef_sim_dc >= coef_obs_dc) + 1) / (nsim + 1)

hist(
  coef_sim_dc,
  main = "Washington, DC: Monte Carlo Distribution of pop_im Coefficients",
  xlab = "Simulated pop_im Coefficient",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(min(coef_sim_dc, coef_obs_dc), max(coef_sim_dc, coef_obs_dc))
)
abline(v = coef_obs_dc, col = "red", lwd = 3)

p_value_dc <- (sum(coef_sim_dc >= coef_obs_dc) + 1) / (nsim + 1)
p_value_dc
```

### New York, NY

```{r}
model_obs_nyc <- ppm(dock_ppp_nyc ~ pop_im_nyc)
coef_obs_nyc <- coef(model_obs_nyc)["pop_im_nyc"]

nsim <- 999
coef_sim_nyc <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_nyc <- runifpoint(n = dock_ppp_nyc$n, win = boundary_owin_nyc)
  sim_model_nyc <- ppm(sim_ppp_nyc ~ pop_im_nyc)
  coef_sim_nyc[i] <- coef(sim_model_nyc)["pop_im_nyc"]
}

p_value_nyc <- (sum(coef_sim_nyc >= coef_obs_nyc) + 1) / (nsim + 1)

hist(
  coef_sim_nyc,
  main = "New York, NY: Monte Carlo Distribution of pop_im Coefficients",
  xlab = "Simulated pop_im Coefficient",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(min(coef_sim_nyc, coef_obs_nyc), max(coef_sim_nyc, coef_obs_nyc))
)
abline(v = coef_obs_nyc, col = "red", lwd = 3)

p_value_nyc <- (sum(coef_sim_nyc >= coef_obs_nyc) + 1) / (nsim + 1)
p_value_nyc
```

### Boston, MA

```{r}
model_obs_bos <- ppm(dock_ppp_bos ~ pop_im_bos)
coef_obs_bos <- coef(model_obs_bos)["pop_im_bos"]

nsim <- 999
coef_sim_bos <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_bos <- runifpoint(n = dock_ppp_bos$n, win = boundary_owin_bos)
  sim_model_bos <- ppm(sim_ppp_bos ~ pop_im_bos)
  coef_sim_bos[i] <- coef(sim_model_bos)["pop_im_bos"]
}

p_value_bos <- (sum(coef_sim_bos >= coef_obs_bos) + 1) / (nsim + 1)

hist(
  coef_sim_bos,
  main = "Boston, MA: Monte Carlo Distribution of pop_im Coefficients",
  xlab = "Simulated pop_im Coefficient",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(min(coef_sim_bos, coef_obs_bos), max(coef_sim_bos, coef_obs_bos))
)
abline(v = coef_obs_bos, col = "red", lwd = 3)

p_value_bos <- (sum(coef_sim_bos >= coef_obs_bos) + 1) / (nsim + 1)
p_value_bos
```

### San Francisco, CA

```{r}
model_obs_sf <- ppm(dock_ppp_sf ~ pop_im_sf)
coef_obs_sf <- coef(model_obs_sf)["pop_im_sf"]

nsim <- 999
coef_sim_sf <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_sf <- runifpoint(n = dock_ppp_sf$n, win = boundary_owin_sf)
  sim_model_sf <- ppm(sim_ppp_sf ~ pop_im_sf)
  coef_sim_sf[i] <- coef(sim_model_sf)["pop_im_sf"]
}

p_value_sf <- (sum(coef_sim_sf >= coef_obs_sf) + 1) / (nsim + 1)

hist(
  coef_sim_sf,
  main = "San Francisco, CA: Monte Carlo Distribution of pop_im Coefficients",
  xlab = "Simulated pop_im Coefficient",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(min(coef_sim_sf, coef_obs_sf), max(coef_sim_sf, coef_obs_sf))
)
abline(v = coef_obs_sf, col = "red", lwd = 3)

p_value_sf <- (sum(coef_sim_sf >= coef_obs_sf) + 1) / (nsim + 1)
p_value_sf
```

### Chicago, IL

```{r}
model_obs_chi <- ppm(dock_ppp_chi ~ pop_im_chi)
coef_obs_chi <- coef(model_obs_chi)["pop_im_chi"]

nsim <- 999
coef_sim_chi <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_chi <- runifpoint(n = dock_ppp_chi$n, win = boundary_owin_chi)
  sim_model_chi <- ppm(sim_ppp_chi ~ pop_im_chi)
  coef_sim_chi[i] <- coef(sim_model_chi)["pop_im_chi"]
}

p_value_chi <- (sum(coef_sim_chi >= coef_obs_chi) + 1) / (nsim + 1)

hist(
  coef_sim_chi,
  main = "Chicago, IL: Monte Carlo Distribution of pop_im Coefficients",
  xlab = "Simulated pop_im Coefficient",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(min(coef_sim_chi, coef_obs_chi), max(coef_sim_chi, coef_obs_chi))
)
abline(v = coef_obs_chi, col = "red", lwd = 3)

p_value_chi <- (sum(coef_sim_chi >= coef_obs_chi) + 1) / (nsim + 1)
p_value_chi
```
:::

### Monte Carlo Simulation Results against Population Density

| Metropolitan Area | p-value |
|-------------------|---------|
| Washington, DC    |         |
| New York, NY      |         |
| Boston, MA        |         |
| San Francisco, CA |         |
| Chicago, IL       |         |

## Hypothesis Test: Other Transit options

### Correlation Functions

::: panel-tabset
### Washington, DC

```{r}
stations_proj_dc <- st_transform(stations_dc, 32618)
docks_proj_dc <- st_transform(docks_dc, 32618)
boundary_proj_dc <- st_transform(hull_buffered_wgs84_dc, 32618)

boundary_owin_dc <- as.owin(boundary_proj_dc)
dock_coords_dc <- st_coordinates(docks_proj_dc)
dock_ppp_dc <- ppp(x = dock_coords_dc[,1],
                y = dock_coords_dc[,2],
                window = boundary_owin_dc)

station_coords_dc <- st_coordinates(stations_proj_dc)
station_ppp_dc <- ppp(x = station_coords_dc[,1],
                   y = station_coords_dc[,2],
                   window = boundary_owin_dc)

dist_im_dc <- distmap(station_ppp_dc)
```

```{r}
m0_dc <- ppm(dock_ppp_dc ~ 1)
m1_dc <- ppm(dock_ppp_dc ~ dist_im_dc)

anova(m0_dc, m1_dc, test = "LR")
summary(m1_dc)
```

### New York, NY

```{r}
stations_proj_nyc <- st_transform(stations_nyc, 32618)
docks_proj_nyc <- st_transform(docks_nyc, 32618)
boundary_proj_nyc <- st_transform(hull_buffered_wgs84_nyc, 32618)

boundary_owin_nyc <- as.owin(boundary_proj_nyc)
dock_coords_nyc <- st_coordinates(docks_proj_nyc)
dock_ppp_nyc <- ppp(x = dock_coords_nyc[,1],
                y = dock_coords_nyc[,2],
                window = boundary_owin_nyc)

station_coords_nyc <- st_coordinates(stations_proj_nyc)
station_ppp_nyc <- ppp(x = station_coords_nyc[,1],
                   y = station_coords_nyc[,2],
                   window = boundary_owin_nyc)

dist_im_nyc <- distmap(station_ppp_nyc)
```

```{r}
m0_nyc <- ppm(dock_ppp_nyc ~ 1)
m1_nyc <- ppm(dock_ppp_nyc ~ dist_im_nyc)

anova(m0_nyc, m1_nyc, test = "LR")
summary(m1_nyc)
```

### Boston, MA

```{r}
stations_proj_bos <- st_transform(stations_bos, 32618)
docks_proj_bos <- st_transform(docks_bos, 32618)
boundary_proj_bos <- st_transform(hull_buffered_wgs84_bos, 32618)

boundary_owin_bos <- as.owin(boundary_proj_bos)
dock_coords_bos <- st_coordinates(docks_proj_bos)
dock_ppp_bos <- ppp(x = dock_coords_bos[,1],
                y = dock_coords_bos[,2],
                window = boundary_owin_bos)

station_coords_bos <- st_coordinates(stations_proj_bos)
station_ppp_bos <- ppp(x = station_coords_bos[,1],
                   y = station_coords_bos[,2],
                   window = boundary_owin_bos)

dist_im_bos <- distmap(station_ppp_bos)
```

```{r}
m0_bos <- ppm(dock_ppp_bos ~ 1)
m1_bos <- ppm(dock_ppp_bos ~ dist_im_bos)

anova(m0_bos, m1_bos, test = "LR")
summary(m1_bos)
```

### San Francisco, CA

```{r}
stations_proj_sf <- st_transform(stations_sf, 32618)
docks_proj_sf <- st_transform(docks_sf, 32618)
boundary_proj_sf <- st_transform(hull_buffered_wgs84_sf, 32618)

boundary_owin_sf <- as.owin(boundary_proj_sf)
dock_coords_sf <- st_coordinates(docks_proj_sf)
dock_ppp_sf <- ppp(x = dock_coords_sf[,1],
                y = dock_coords_sf[,2],
                window = boundary_owin_sf)

station_coords_sf <- st_coordinates(stations_proj_sf)
station_ppp_sf <- ppp(x = station_coords_sf[,1],
                   y = station_coords_sf[,2],
                   window = boundary_owin_sf)

dist_im_sf <- distmap(station_ppp_sf)
```

```{r}
m0_sf <- ppm(dock_ppp_sf ~ 1)
m1_sf <- ppm(dock_ppp_sf ~ dist_im_sf)

anova(m0_sf, m1_sf, test = "LR")
summary(m1_sf)
```

### Chicago, IL

```{r}
stations_proj_chi <- st_transform(stations_chi, 32618)
docks_proj_chi <- st_transform(docks_chi, 32618)
boundary_proj_chi <- st_transform(hull_buffered_wgs84_chi, 32618)

boundary_owin_chi <- as.owin(boundary_proj_chi)
dock_coords_chi <- st_coordinates(docks_proj_chi)
dock_ppp_chi <- ppp(x = dock_coords_chi[,1],
                y = dock_coords_chi[,2],
                window = boundary_owin_chi)

station_coords_chi <- st_coordinates(stations_proj_chi)
station_ppp_chi <- ppp(x = station_coords_chi[,1],
                   y = station_coords_chi[,2],
                   window = boundary_owin_chi)

dist_im_chi <- distmap(station_ppp_chi)
```

```{r}
m0_chi <- ppm(dock_ppp_chi ~ 1)
m1_chi <- ppm(dock_ppp_chi ~ dist_im_chi)

anova(m0_chi, m1_chi, test = "LR")
summary(m1_chi)
```
:::

### Relationship to Train Station Locations (Poisson Point Process)

| Metropolitan Area | Docks | Dock Intensity (points/meter squared) | `dist_im` coefficient | Z-value | Deviance diff |
|------------|------------|------------|------------|------------|------------|
| Washington, DC | $774$ | $5.06 \times 10^{-7}$ | $-1.09 \times 10^{-3}$ | $-23.96$ | $1112.3$ |
| New York, NY | $2,303$ | $1.5 \times 10^{-5}$ | $-1.61 \times 10^{-3}$ | $-25.90$ | $1174.6$ |
| Boston, MA | $320$ | $1.08 \times 10^{-5}$ | $-1.46 \times 10^{-3}$ | $-12.29$ | $263.95$ |
| San Francisco, CA | $107$ | $8.44 \times 10^{-7}$ | $-2.97 \times 10^{-04}$ | $-2.83$ | $9.655$ |
| Chicago, IL | $909$ | $3.77 \times 10^{-6}$ | $-8.24 \times 10^{-4}$ | $-17.02$ | $386.02$ |

### Monte Carlo Simulations

::: panel-tabset
### Washington, DC

```{r}
stations_proj_dc <- st_transform(stations_dc, 32618)
station_coords_dc <- st_coordinates(stations_proj_dc)

station_ppp_dc <- ppp(
  x = station_coords_dc[,1],
  y = station_coords_dc[,2],
  window = boundary_owin_dc
)

obs_dist_dc <- nndist(dock_ppp_dc, station_ppp_dc)
obs_mean_dist_dc <- mean(obs_dist_dc)


nsim <- 999
sim_mean_dist_dc <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_dc <- rpoint(
    n = dock_ppp_dc$n,
    f = pop_im_dc,
    win = boundary_owin_dc
  )
  sim_dist <- nndist(sim_ppp_dc, station_ppp_dc)
  sim_mean_dist_dc[i] <- mean(sim_dist)
}


p_value_dc <- (sum(sim_mean_dist_dc <= obs_mean_dist_dc) + 1) / (nsim + 1)

x_min <- min(sim_mean_dist_dc, obs_mean_dist_dc)
x_max <- max(sim_mean_dist_dc, obs_mean_dist_dc)

hist(
  sim_mean_dist_dc,
  main = "Washington, DC: Monte Carlo Distribution of\nDistance from Docks to Nearest Train Station",
  xlab = "Simulated Mean Distance",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(x_min, x_max)
)

abline(v = obs_mean_dist_dc, col = "red", lwd = 3)
p_value_dc
```

### New York, NY

```{r}
stations_proj_nyc <- st_transform(stations_nyc, 32618)
station_coords_nyc <- st_coordinates(stations_proj_nyc)

station_ppp_nyc <- ppp(
  x = station_coords_nyc[,1],
  y = station_coords_nyc[,2],
  window = boundary_owin_nyc
)

obs_dist_nyc <- nndist(dock_ppp_nyc, station_ppp_nyc)
obs_mean_dist_nyc <- mean(obs_dist_nyc)


nsim <- 999
sim_mean_dist_nyc <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_nyc <- rpoint(
    n = dock_ppp_nyc$n,
    f = pop_im_nyc,
    win = boundary_owin_nyc
  )
  sim_dist <- nndist(sim_ppp_nyc, station_ppp_nyc)
  sim_mean_dist_nyc[i] <- mean(sim_dist)
}


p_value_nyc <- (sum(sim_mean_dist_nyc <= obs_mean_dist_nyc) + 1) / (nsim + 1)

x_min <- min(sim_mean_dist_nyc, obs_mean_dist_nyc)
x_max <- max(sim_mean_dist_nyc, obs_mean_dist_nyc)

hist(
  sim_mean_dist_nyc,
  main = "New York, NY: Monte Carlo Distribution of\nDistance from Docks to Nearest Train Station",
  xlab = "Simulated Mean Distance",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(x_min, x_max)
)

abline(v = obs_mean_dist_nyc, col = "red", lwd = 3)
p_value_nyc
```

### Boston, MA

```{r}
stations_proj_bos <- st_transform(stations_bos, 32618)
station_coords_bos <- st_coordinates(stations_proj_bos)

station_ppp_bos <- ppp(
  x = station_coords_bos[,1],
  y = station_coords_bos[,2],
  window = boundary_owin_bos
)

obs_dist_bos <- nndist(dock_ppp_bos, station_ppp_bos)
obs_mean_dist_bos <- mean(obs_dist_bos)


nsim <- 999
sim_mean_dist_bos <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_bos <- rpoint(
    n = dock_ppp_bos$n,
    f = pop_im_bos,
    win = boundary_owin_bos
  )
  sim_dist <- nndist(sim_ppp_bos, station_ppp_bos)
  sim_mean_dist_bos[i] <- mean(sim_dist)
}


p_value_bos <- (sum(sim_mean_dist_bos <= obs_mean_dist_bos) + 1) / (nsim + 1)

x_min <- min(sim_mean_dist_bos, obs_mean_dist_bos)
x_max <- max(sim_mean_dist_bos, obs_mean_dist_bos)

hist(
  sim_mean_dist_bos,
  main = "Boston, MA: Monte Carlo Distribution of\nDistance from Docks to Nearest Train Station",
  xlab = "Simulated Mean Distance",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(x_min, x_max)
)

abline(v = obs_mean_dist_bos, col = "red", lwd = 3)
p_value_bos
```

### San Francisco, CA

```{r}
stations_proj_sf <- st_transform(stations_sf, 32618)
station_coords_sf <- st_coordinates(stations_proj_sf)

station_ppp_sf <- ppp(
  x = station_coords_sf[,1],
  y = station_coords_sf[,2],
  window = boundary_owin_sf
)

obs_dist_sf <- nndist(dock_ppp_sf, station_ppp_sf)
obs_mean_dist_sf <- mean(obs_dist_sf)


nsim <- 999
sim_mean_dist_sf <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_sf <- rpoint(
    n = dock_ppp_sf$n,
    f = pop_im_sf,
    win = boundary_owin_sf
  )
  sim_dist <- nndist(sim_ppp_sf, station_ppp_sf)
  sim_mean_dist_sf[i] <- mean(sim_dist)
}


p_value_sf <- (sum(sim_mean_dist_sf <= obs_mean_dist_sf) + 1) / (nsim + 1)

x_min <- min(sim_mean_dist_sf, obs_mean_dist_sf)
x_max <- max(sim_mean_dist_sf, obs_mean_dist_sf)

hist(
  sim_mean_dist_sf,
  main = "San Francisco, CA: Monte Carlo Distribution of\nDistance from Docks to Nearest Train Station",
  xlab = "Simulated Mean Distance",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(x_min, x_max)
)

abline(v = obs_mean_dist_sf, col = "red", lwd = 3)
p_value_sf
```

### Chicago, IL

```{r}
stations_proj_chi <- st_transform(stations_chi, 32618)
station_coords_chi <- st_coordinates(stations_proj_chi)

station_ppp_chi <- ppp(
  x = station_coords_chi[,1],
  y = station_coords_chi[,2],
  window = boundary_owin_chi
)

obs_dist_chi <- nndist(dock_ppp_chi, station_ppp_chi)
obs_mean_dist_chi <- mean(obs_dist_chi)


nsim <- 999
sim_mean_dist_chi <- numeric(nsim)

for (i in 1:nsim) {
  sim_ppp_chi <- rpoint(
    n = dock_ppp_chi$n,
    f = pop_im_chi,
    win = boundary_owin_chi
  )
  sim_dist <- nndist(sim_ppp_chi, station_ppp_chi)
  sim_mean_dist_chi[i] <- mean(sim_dist)
}


p_value_chi <- (sum(sim_mean_dist_chi <= obs_mean_dist_chi) + 1) / (nsim + 1)

x_min <- min(sim_mean_dist_chi, obs_mean_dist_chi)
x_max <- max(sim_mean_dist_chi, obs_mean_dist_chi)

hist(
  sim_mean_dist_chi,
  main = "Chicago, IL: Monte Carlo Distribution of\nDistance from Docks to Nearest Train Station",
  xlab = "Simulated Mean Distance",
  breaks = 30,
  col = "lightgray",
  border = "white",
  xlim = c(x_min, x_max)
)

abline(v = obs_mean_dist_chi, col = "red", lwd = 3)
p_value_chi
```
:::

### Monte Carlo Simulation Results against Train Stations

| Metropolitan Area | p-value |
|-------------------|---------|
| Washington, DC    |         |
| New York, NY      |         |
| Boston, MA        |         |
| San Francisco, CA |         |
| Chicago, IL       |         |

## Discussion